---
title: "Epiparam 2.0"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    navbar:
    - icon: "fa fa-rss"
      href: http://onehealth.or.kr/
      align: right
    vertical_layout: fill
    theme:
      version: 4
      fg: '#030303'
      bg: '#FBFBFB'
      navbar-bg: '#8B5A00'
      sidebar-bg: '#8B5A00'
      smart: no
runtime: shiny
resource_files:
- gamma_model.rds
- lognormal_model.rds
- normal_model.rds
- weibull_model.rds
---

```{=html}
<style type="text/css">

.vis-network .vis-export {
position: absolute !important;
top: 10px !important;
right: 10px !important;
z-index: 1000 !important;
background-color: #4169E1 !important;
color: white !important;
border: none !important;
border-radius: 4px !important;
padding: 6px 12px !important;
font-weight: bold !important;
margin-right: 0 !important;
}

.vis-network .vis-export:hover {
background-color: #3151A3 !important;
}

.irs--shiny .irs-min, .irs--shiny .irs-max {
top: 0;
padding: 1px 3px;
text-shadow: none;
background-color: rgba(25, 25, 26, 0.1);
border-radius: 3px;
font-size: 14px;
line-height: 1.333;
}

.irs--shiny .irs-from, .irs--shiny .irs-to, .irs--shiny .irs-single {
color: #FBFBFB;
text-shadow: none;
padding: 1px 3px;
background-color: #000000;
border-radius: 3px;
font-size: 14px;
line-height: 1.333;
}

.irs--shiny .irs-bar {
top: 25px;
height: 8px;
border-top: 1px solid #70b2e0;
border-bottom: 1px solid #70b2e0;
background: #70b2e0;
cursor: s-resize;
z-index: 2;
}

.error-highlight {
border: 2px solid #ff0000 !important;
background-color: #ffeeee !important;
}

body hr {
border-color: #ffffff;
margin: 10px 0;
border: 2px solid #ffffff !important;
}

.rt-bordered .rt-table, .rt-outlined .rt-table {
border-style: solid;
border-top-left-radius: 1.25rem;
}

h4, .h4 {
font-size: 10px;
}

.chart-wrapper .chart-stage {
color: #000000;
padding: 0px;
}

.navbar.navbar-inverse {
background-color: #8B5A00 !important;
border-bottom-right-radius: 5.25rem;
}

.navbar-brand, .navbar.navbar-inverse .navbar-brand {
color: white !important;
}

.navbar-brand:hover, .navbar.navbar-inverse .navbar-brand:hover {
color: white !important;
}

.nav-tabs-custom > .nav-tabs > li.active {
border-top-color: #8B5A00;
}

.navbar.navbar-inverse ul.nav.navbar-nav>li {
color: #fff;
font-size: 16px;
font-family: "TT Norms Pro";
}

.nav-tabs-custom > .nav-tabs {
margin: 0;
border-bottom: 1px solid #8B5A00;
color: white;
border-top-right-radius: 10px;
border-top-left-radius: 3px;
font-family: "Tw Cen MT";
}

.nav-tabs-custom > .nav-tabs > li > a, .nav-tabs-custom > .nav-tabs > li > a:active {
color: #000000;
font-size: 0.9375rem;
border-radius: 0;
padding: 3px 10px 5px;
text-transform: none;
font-weight: 500;
}


.chart-wrapper, .nav-tabs-custom, .sbframe-commentary {
background: #FFFFFF;
border: 1px solid #B2B2B2;
border-radius: 0.25rem;
margin-bottom: 8px;
margin-right: 8px;
}

.navbar-brand {
font-family: "TT Norms Pro Black";
font-size: 38px;
padding: 0px;
}

.chart-title {
border-bottom: 1px solid #B7B7B7;
font-size: 0.9375rem;
text-align: center;
font-weight: bold;
padding: 7px 10px 4px;
}

.section.sidebar {
font-family: "TT Norms Pro";
font-size: 14px;
background-color: #8B5A00 !important;
color: white;
padding: 0px;
border: 1px solid #8B5A00;
position: fixed;
top: 62px;
left: 0;
bottom: 0;
border-right: 1px solid #dee2e6;
background-color: rgba(39, 128, 227, 0.1);
padding-top: 10px;
padding-left: 10px;
padding-right: 10px;
visibility: hidden;
overflow: auto;
}

a {
color: #28282b;
font-size: 19px;
font-family: TT Norms Pro;
background-color: transparent;
}

.nav-tabs .nav-link.active, .nav-tabs>li>a.active, .nav-tabs .nav-pills>li>a.active, .nav-tabs :where(ul.nav.navbar-nav > li)>a.active, .nav-tabs .nav-item.show .nav-link, .nav-tabs .nav-item.in .nav-link, .nav-tabs .nav-item.show .nav-tabs>li>a, .nav-tabs .nav-item.in .nav-tabs>li>a, .nav-tabs .nav-item.show .nav-pills>li>a, .nav-tabs .nav-item.in .nav-pills>li>a, .nav-tabs>li.show .nav-link, .nav-tabs>li.in .nav-link, .nav-tabs>li.show .nav-tabs>li>a, .nav-tabs>li.in .nav-tabs>li>a, .nav-tabs>li.show .nav-pills>li>a, .nav-tabs>li.in .nav-pills>li>a, .nav-tabs .nav-pills>li.show .nav-link, .nav-tabs .nav-pills>li.in .nav-link, .nav-tabs .nav-pills>li.show .nav-tabs>li>a, .nav-tabs .nav-pills>li.in .nav-tabs>li>a, .nav-tabs .nav-pills>li.show .nav-pills>li>a, .nav-tabs .nav-pills>li.in .nav-pills>li>a, .nav-tabs .nav-item.show :where(ul.nav.navbar-nav > li)>a, .nav-tabs .nav-item.in :where(ul.nav.navbar-nav > li)>a, .nav-tabs>li.show :where(ul.nav.navbar-nav > li)>a, .nav-tabs>li.in :where(ul.nav.navbar-nav > li)>a, .nav-tabs .nav-pills>li.show :where(ul.nav.navbar-nav > li)>a, .nav-tabs .nav-pills>li.in :where(ul.nav.navbar-nav > li)>a, .nav-tabs .show:where(ul.nav.navbar-nav > li):not(.dropdown) .nav-link, .nav-tabs .in:where(ul.nav.navbar-nav > li):not(.dropdown) .nav-link, .nav-tabs .show:where(ul.nav.navbar-nav > li):not(.dropdown) .nav-tabs>li>a, .nav-tabs .in:where(ul.nav.navbar-nav > li):not(.dropdown) .nav-tabs>li>a, .nav-tabs .show:where(ul.nav.navbar-nav > li):not(.dropdown) .nav-pills>li>a, .nav-tabs .in:where(ul.nav.navbar-nav > li):not(.dropdown) .nav-pills>li>a, .nav-tabs .show:where(ul.nav.navbar-nav > li):not(.dropdown) :where(ul.nav.navbar-nav > li)>a, .nav-tabs .in:where(ul.nav.navbar-nav > li):not(.dropdown) :where(ul.nav.navbar-nav > li)>a {
color: #b814aa;
background-color: #FBFBFB;
border-color: #B7B7B7 #B7B7B7 #FBFBFB;
}

.chart-wrapper {
width: 100% !important;
}

pre {
font-family: "TT Norms Pro";
font-size: medium;
}

.btn-primary {
color: #FBFBFB;
background-color: #ff7100;
border-color: #ff7100;
}


</style>

```

```{r setup-chunk, include=FALSE}
# load libraries
library(shinyjs)
library(htmltools)
library(knitr)
library(flexdashboard)
library(lubridate)
library(bslib)
library(shiny)
library(tidyverse)
library(janitor)
library(plotly)
library(MASS)
library(R0)
library(DT)
library(EpiEstim)
library(incidence2)
library(visNetwork)
library(webshot)
library(webshot2)
library(officer)
library(networkD3)
library(highcharter)
library(fitdistrplus)
library(stringr)
library(rstan)
library(loo)

# suppress browser redirect on Stan warnings
options(warn = -1) 
rstan_options(auto_write = TRUE)

# date parsing function
parse_date <- function(x) {
  x <- trimws(as.character(x))
  parsed <- lubridate::parse_date_time(x, orders = c("dmy", "ymd", "mdy"), quiet = TRUE)
  as.Date(parsed)
}

# helper functions for model fitting (MLE pages)
fit_parametric_model <- function(data, dist, shift_val = 0) {
  data_fit <- data + shift_val
  
  fit <- tryCatch({
    MASS::fitdistr(data_fit, dist)
  }, error = function(e) return(NULL))
  
  if(is.null(fit)) return(NULL)
  
  est <- fit$estimate
  
  if (dist == "lognormal") {
    mean_val <- exp(est["meanlog"] + est["sdlog"]^2 / 2)
    var_val <- (exp(est["sdlog"]^2) - 1) * exp(2 * est["meanlog"] + est["sdlog"]^2)
    sd_val <- sqrt(var_val)
    ci_95 <- qlnorm(0.95, est["meanlog"], est["sdlog"])
    
    raw_params <- data.frame(
      parameter = c("estimate.meanlog", "estimate.sdlog", "sd.meanlog", "sd.sdlog"),
      value = round(c(est["meanlog"], est["sdlog"], fit$sd["meanlog"], fit$sd["sdlog"]), 2)
    )
    
  } else if (dist == "weibull") {
    mean_val <- est["scale"] * gamma(1 + 1/est["shape"])
    var_val <- est["scale"]^2 * (gamma(1 + 2/est["shape"]) - (gamma(1 + 1/est["shape"]))^2)
    sd_val <- sqrt(var_val)
    ci_95 <- qweibull(0.95, est["shape"], est["scale"])
    
    raw_params <- data.frame(
      parameter = c("estimate.shape", "estimate.scale", "sd.shape", "sd.scale"),
      value = round(c(est["shape"], est["scale"], fit$sd["shape"], fit$sd["scale"]), 2)
    )
    
  } else if (dist == "gamma") {
    mean_val <- est["shape"] / est["rate"]
    var_val <- est["shape"] / (est["rate"]^2)
    sd_val <- sqrt(var_val)
    ci_95 <- qgamma(0.95, est["shape"], est["rate"])
    
    raw_params <- data.frame(
      parameter = c("estimate.shape", "estimate.rate", "sd.shape", "sd.rate"),
      value = round(c(est["shape"], est["rate"], fit$sd["shape"], fit$sd["rate"]), 2)
    )
    
  } else if (dist == "normal") {
    mean_val <- est["mean"]
    sd_val <- est["sd"]
    ci_95 <- qnorm(0.95, mean_val, sd_val)
    var_val <- sd_val^2
    
    raw_params <- data.frame(
      parameter = c("estimate.mean", "estimate.sd", "sd.mean", "sd.sd"),
      value = round(c(est["mean"], est["sd"], fit$sd["mean"], fit$sd["sd"]), 2)
    )
  }
  
  final_mean <- as.numeric(mean_val) - shift_val
  final_var <- as.numeric(var_val)
  final_ci <- as.numeric(ci_95) - shift_val
  
  list(
    model = toupper(dist),
    fit_obj = fit,
    raw_params_df = raw_params,
    moments = c(mean=final_mean, var=final_var, ci95=final_ci),
    aic = AIC(fit),
    bic = BIC(fit)
  )
}

# helper to generate plot data for incubation (MLE page)
get_plot_data_inc <- function(data, fit_res) {
  h_breaks <- seq(floor(min(data))-0.5, ceiling(max(data))+0.5, by = 1)
  hist_data <- hist(data, plot = FALSE, breaks = h_breaks)
  
  hist_df <- data.frame(
    from = hist_data$breaks[-length(hist_data$breaks)],
    to = hist_data$breaks[-1],
    y = hist_data$density,
    x = hist_data$mids
  )
  
  est <- fit_res$fit_obj$estimate
  d_name <- tolower(fit_res$model)
  min_x_calc <- max(0.01, min(data) - 1.5)
  x_vals <- seq(min_x_calc, max(data) + 1.5, length.out = 500)
  
  suppressWarnings({
    y_vals <- tryCatch({
      switch(d_name,
             "lognormal" = dlnorm(x_vals, meanlog = est["meanlog"], sdlog = est["sdlog"]),
             "weibull" = dweibull(x_vals, shape = est["shape"], scale = est["scale"]),
             "gamma"  = dgamma(x_vals, shape = est["shape"], rate = est["rate"])
      )
    }, error = function(e) return(rep(0, length(x_vals))))
  })
  
  curve_data <- data.frame(x = x_vals, y = y_vals)
  list(hist = hist_df, curve = curve_data)
}

# helper to generate plot data for serial interval (MLE page)
get_plot_data_si <- function(data, fit_res) {
  h_breaks <- seq(floor(min(data))-1, ceiling(max(data))+1, by = 2) 
  hist_data <- hist(data, plot = FALSE, breaks = h_breaks)
  hist_df <- data.frame(from = hist_data$breaks[-length(hist_data$breaks)], to = hist_data$breaks[-1], y = hist_data$density, x = hist_data$mids)
  
  shift_val <- if(tolower(fit_res$model) == "normal") 0 else 1
  est <- fit_res$fit_obj$estimate
  d_name <- tolower(fit_res$model)
  x_vals <- seq(min(data) - 5, max(data) + 5, length.out = 500)
  x_trans <- x_vals + shift_val
  
  if(shift_val > 0) x_trans[x_trans <= 0] <- NA
  
  suppressWarnings({
    y_vals <- tryCatch({
      switch(d_name,
             "lognormal" = dlnorm(x_trans, meanlog = est["meanlog"], sdlog = est["sdlog"]),
             "weibull"   = dweibull(x_trans, shape = est["shape"], scale = est["scale"]),
             "gamma"     = dgamma(x_trans, shape = est["shape"], rate = est["rate"]),
             "normal"    = dnorm(x_trans, mean = est["mean"], sd = est["sd"])
      )
    }, error = function(e) return(rep(0, length(x_vals))))
  })
  y_vals[is.na(y_vals)] <- 0
  curve_data <- data.frame(x = x_vals, y = y_vals)
  list(hist = hist_df, curve = curve_data)
}

# bayesian helper: get stan code (not actually used but for reference)
get_stan_code <- function(model_name = "Weibull") {
  if (model_name == "Weibull") {
    return("
      data { int<lower=0> N; vector[N] tStartExposure; vector[N] tEndExposure; vector[N] tSymptomOnset; }
      parameters { real<lower=0> alphaInc; real<lower=0> sigmaInc; vector<lower=0, upper=1>[N] uE; }
      transformed parameters { vector[N] tE; tE = tStartExposure + uE .* (tEndExposure - tStartExposure); }
      model { alphaInc ~ normal(0, 10); sigmaInc ~ normal(0, 10); target += weibull_lpdf(tSymptomOnset - tE | alphaInc, sigmaInc); }
    ")
  } 
  else if (model_name == "Gamma") {
    return("
      data { int<lower=0> N; vector[N] tStartExposure; vector[N] tEndExposure; vector[N] tSymptomOnset; }
      parameters { real<lower=0> alphaInc; real<lower=0> sigmaInc; vector<lower=0, upper=1>[N] uE; }
      transformed parameters { vector[N] tE; tE = tStartExposure + uE .* (tEndExposure - tStartExposure); }
      model { alphaInc ~ normal(0, 10); sigmaInc ~ normal(0, 10); target += gamma_lpdf(tSymptomOnset - tE | alphaInc, sigmaInc); }
    ")
  }
  else if (model_name == "Lognormal") {
    return("
      data { int<lower=0> N; vector[N] tStartExposure; vector[N] tEndExposure; vector[N] tSymptomOnset; }
      parameters { real alphaInc; real<lower=0> sigmaInc; vector<lower=0, upper=1>[N] uE; }
      transformed parameters { vector[N] tE; tE = tStartExposure + uE .* (tEndExposure - tStartExposure); }
      model { alphaInc ~ normal(0, 10); sigmaInc ~ normal(0, 10); target += lognormal_lpdf(tSymptomOnset - tE | alphaInc, sigmaInc); }
    ")
  }
  else if (model_name == "Normal") {
    return("
      data { int<lower=0> N; vector[N] tStartExposure; vector[N] tEndExposure; vector[N] tSymptomOnset; }
      parameters { real alphaInc; real<lower=0> sigmaInc; vector<lower=0, upper=1>[N] uE; }
      transformed parameters { vector[N] tE; tE = tStartExposure + uE .* (tEndExposure - tStartExposure); }
      model { alphaInc ~ normal(0, 10); sigmaInc ~ normal(0, 10); target += normal_lpdf(tSymptomOnset - tE | alphaInc, sigmaInc); }
    ")
  }
  return(NULL)
}

# load stan models
get_compiled_stan_model <- function(model_name = "Weibull") {
  
  # object name for in-memory caching
  obj_name <- paste0("stan_model_", model_name)
  file_name <- paste0(tolower(model_name), "_model.rds")
  
  # check memory (RAM)
  if (exists(obj_name, envir = .GlobalEnv)) {
    return(get(obj_name, envir = .GlobalEnv))
  }
  
  # check disk (load the Linux/windows binary)
  if (file.exists(file_name)) {
    # this reads the binary we made in WSL/Windows
    mod <- readRDS(file_name) 
    assign(obj_name, mod, envir = .GlobalEnv)
    return(mod)
  }
  
  stop(paste("Model file", file_name, "not found. Ensure .rds files are uploaded."))
}

```

```{r global-downloads, context="server"}

# global mock data handlers
output$downloadMockIncubation <- downloadHandler(
  filename = function() { paste0("mock_incubation-", Sys.Date(), ".csv") },
  content = function(file) { write.csv(read.csv("Busan_real.csv"), file) }
)

output$downloadMockserial <- downloadHandler(
  filename = function() { paste0("mock_serial-", Sys.Date(), ".csv") },
  content = function(file) { write.csv(read.csv("Serial.csv"), file) }
)

output$downloadMockKappa <- downloadHandler(
  filename = function() { paste0("mock_kappa_data-", Sys.Date(), ".csv") },
  content = function(file) { write.csv(read.csv("kappa_data.csv"), file) }
)

output$downloadMockRt <- downloadHandler(
  filename = function() { paste0("mock_Rt_data-", Sys.Date(), ".csv") },
  content = function(file) { write.csv(read.csv("rt.csv"), file, row.names = FALSE) }
)

output$downloadMockIncubationBayes <- downloadHandler(
  filename = function() { paste0("mock_incubation_bayesian-", Sys.Date(), ".csv") },
  content = function(file) { 
    write.csv(read.csv("incubation_bayesian.csv"), file, row.names = FALSE) 
  }
)

output$downloadMockSerialBayes <- downloadHandler(
  filename = function() { paste0("mock_serial_bayesian-", Sys.Date(), ".csv") },
  content = function(file) { 
    write.csv(read.csv("serial_bayesian.csv"), file, row.names = FALSE) 
  }
)

output$downloadUserManual <- downloadHandler(
  filename = "User_manual.pdf",
  content = function(file){
    file.copy("User_manual.pdf", file)
  }
)

```

Incubation Period {data-orientation=rows}
==========================================

Sidebar {.sidebar}
--------------------

```{r inc-sidebar}

useShinyjs(rmd = TRUE)

hr()

fileInput("file1_inc", "Choose CSV File",
          multiple = TRUE,
          accept = c("text/csv", "text/comma-separated-values,text/plain", ".csv"))
hr()
selectInput(inputId = "sex_inc", label = "Sex",
            choices = c("All", "Male", "Female"),
            selected = "All")

selectInput(inputId = "age_group_inc", label = "Age Group",
            choices = c("All", '0 - 19', '20 - 29', '30 - 39', '40 - 49', '50 - 59', '≥ 60'),
            selected = "All")
hr()

selectInput(inputId = "model_inc", label = "Distribution Model",
            choices = c("Lognormal", "Weibull", "Gamma", "ALL"),
            selected = "Lognormal")

# comparison view toggle for all models
conditionalPanel(
  condition = "input.model_inc == 'ALL'",
  radioButtons("view_inc", "Comparison View:",
               choices = c("Side-by-Side", "Overlaid"),
               selected = "Side-by-Side")
)

hr()
valueBox(
  value = tags$h6("Note: The Incubation Period is estimated using Maximum Likelihood Estimation (MLE).",
                  style = "text-align: justify; font-family: Tw Cen MT; font-size: 20px; color: white;"),
  color = "#FFFFFF")

hr()
valueBox(
  value = downloadButton("downloadMockIncubation", "Download Mock Data", icon = icon("download")),
  color = "#FFFFFF"
)

hr()
valueBox(
  value = downloadButton("downloadUserManual", "Download User Manual", icon = icon("download")),
  color = "#FFFFFF"
)

renderImage({
  pfad <- "one_logo.png"
  list(
    src = pfad,
    contentType = 'image/png',
    width = 210,
    height = 60,
    style = "position: fixed; bottom: 10px; left: 10px; cursor: pointer;",
    onclick = "window.location.href = '[http://onehealth.or.kr/](http://onehealth.or.kr/)';"
  )
}, deleteFile = F)

```

```{r inc-data-processing}

# incubation data processing pipeline

dataset_inc <- eventReactive(list(input$file1_inc, input$sex_inc, input$age_group_inc), {
  req(input$file1_inc)
  
  dataset <- read.csv(input$file1_inc$datapath)
  dataset <- data.frame(lapply(dataset, trimws), stringsAsFactors = FALSE)
  dataset <- clean_names(dataset)
  
  # validation start
  validate(
    need(all(c("date_of_exposure", "date_of_onset") %in% names(dataset)),
         "incorrect dataset: please upload the incubation data containing 'date of exposure' and 'date of onset'.")
  )
  # validation end
  
  if("sex" %in% names(dataset)) {
    dataset <- dataset %>%
      mutate(sex = case_when(
        grepl("^m", sex, ignore.case = TRUE) ~ "Male",
        grepl("^f", sex, ignore.case = TRUE) ~ "Female",
        TRUE ~ sex
      ))
  }
  
  if("age_group" %in% names(dataset)) {
    dataset$age_group <- dataset$age_group %>%
      trimws() %>%              
      gsub("s$", "", ., ignore.case = TRUE) %>%
      gsub("[^0-9]", "", .)
    dataset$age_group <- suppressWarnings(as.numeric(dataset$age_group))
    dataset <- dataset %>% mutate(age_group = case_when(
      age_group > 0 & age_group <= 19 ~ '0 - 19',
      age_group >= 20 & age_group <= 29 ~ '20 - 29',
      age_group >= 30 & age_group <= 39 ~ '30 - 39',
      age_group >= 40 & age_group <= 49 ~ '40 - 49',
      age_group >= 50 & age_group <= 59 ~ '50 - 59',
      TRUE ~ '≥ 60'
    ))
  }
  
  if (input$sex_inc != "All") dataset <- dataset %>% filter(sex == input$sex_inc)
  if (input$age_group_inc != "All") dataset <- dataset %>% filter(age_group == input$age_group_inc)
  
  dataset
})

get_incu_data <- function() {
  req(dataset_inc())
  
  data <- dataset_inc() %>%
    mutate(
      date_of_exposure = parse_date(date_of_exposure),
      date_of_onset  = parse_date(date_of_onset)
    ) %>%
    filter(
      !is.na(date_of_exposure),
      !is.na(date_of_onset)
    ) %>%
    mutate(incuPeriods = as.numeric(date_of_onset - date_of_exposure)) %>%
    filter(incuPeriods > 0) %>%
    pull(incuPeriods) %>%
    na.omit()
  
  if(length(data) == 0) return(NULL)
  data
}

# master fitting logic for incubation
fit_results_inc <- reactive({
  data <- get_incu_data()
  if (is.null(data) || length(data) < 3) return(NULL)
  
  models <- if(input$model_inc == "ALL") c("lognormal", "weibull", "gamma") else tolower(input$model_inc)
  
  res <- list()
  for(m in models) {
    # fit logic
    suppressWarnings({
      fit <- fit_parametric_model(data, m, shift_val = 0)
    })
    if(!is.null(fit)) res[[m]] <- fit
  }
  res
})

```

## Row {data-height="180"}

### Sample Size

```{r inc-sample-size}

renderUI({
  # ensure dataset is available
  req(dataset_inc())
  # data preparation
  ss <- nrow(dataset_inc())
  # render the sample size text
  htmltools::div(
    style = "display: flex; flex-direction: column; justify-content: center; align-items: center; height: 100%; width: 100%; padding: 10px;",
    htmltools::div(
      style = "font-size: 30px; font-weight: bold; margin-bottom: 10px; text-align: center;font-family: Tw Cen MT;",
      ss
    )
  )
})

```

### Model Fit Statistics

```{r inc-fit-stats}

renderDT({
  res <- fit_results_inc()
  data <- get_incu_data()
  
  if (!is.null(data) && length(data) < 3) {
    df <- data.frame(Status = paste("Insufficient data points (N=", length(data), "). Need 3+ incubation period data points."))
  } else if (is.null(res) || length(res) == 0) {
    df <- data.frame(Status = "No data or fit failed")
  } else {
    if (input$model_inc != "ALL") {
      df <- res[[1]]$raw_params_df
    } else {
      df <- do.call(rbind, lapply(res, function(x) {
        data.frame(
          Model = x$model, 
          AIC = round(x$aic, 2), 
          BIC = round(x$bic, 2)
        )
      }))
      df <- df[order(df$AIC), ] 
    }
  }
  
  datatable(
    df, rownames = FALSE,
    options = list(
      dom = "Brt",
      buttons = list("csv", "excel", "pdf"),
      pageLength = nrow(df)
    ),
    extensions = "Buttons"
  ) %>% 
    formatStyle(columns = names(df), `font-size` = "12px") %>%
    formatStyle(columns = intersect(names(df), c("parameter", "Model")), fontWeight = "bold")
})

```

### Model Parameters

```{r inc-model-params}

renderDT({
  res <- fit_results_inc()
  data <- get_incu_data()
  
  if (!is.null(data) && length(data) < 3) {
    df <- data.frame(Status = "Insufficient Data")
  } else if (is.null(res) || length(res) == 0) {
    df <- data.frame(Status = "No data")
  } else {
    fmt_moments <- function(x) {
      data.frame(
        parameter = c(paste(x$model, "Mean"), paste(x$model, "Variance"), "95th Percentile"),
        value = round(c(x$moments["mean"], x$moments["var"], x$moments["ci95"]), 3)
      )
    }
    
    if (input$model_inc != "ALL") {
      df <- fmt_moments(res[[1]])
    } else {
      df <- do.call(rbind, lapply(res, function(x) {
        data.frame(
          Model = x$model,
          `Mean` = round(x$moments["mean"], 3),
          `Variance` = round(x$moments["var"], 3),
          `95th Percentile` = round(x$moments["ci95"], 3),
          check.names = FALSE
        )
      }))
    }
  }
  
  datatable(
    df, rownames = FALSE,
    options = list(
      dom = "Brt",
      buttons = list("csv", "excel", "pdf"),
      pageLength = nrow(df)
    ),
    extensions = "Buttons"
  ) %>% 
    formatStyle(columns = names(df), `font-size` = "12px") %>%
    formatStyle(columns = intersect(names(df), c("parameter", "Model")), fontWeight = "bold")
})

```

Column {.tabset .tabset-fade }
-------------------------------

### Distribution Plots

```{r inc-histograms}

renderHighchart({
  data <- get_incu_data()
  res <- fit_results_inc()
  
  if (is.null(data) || length(data) < 3 || is.null(res) || length(res) == 0) {
    return(highchart() %>% hc_title(text="Insufficient Data (N < 3)"))
  }
  
  col_map <- list(lognormal="#00BFFF", weibull="#FF8C00", gamma="#228B22")
  line_col_overlaid <- list(lognormal="#000000", weibull="#FF0000", gamma="#008000") 
  
  # style definitions
  axis_style <- list(color = "#000000", fontSize = "16px", fontWeight = "bold")
  label_style <- list(color = "#000000", fontSize = "14px", fontWeight = "bold")
  menu_items <- list("viewFullscreen", "separator", "downloadPNG", "downloadPDF", "downloadSVG")
  
  if (input$model_inc == "ALL") {
    
    # calculate global max y for constant scale
    max_y <- 0
    all_plot_data <- list()
    for(m in names(res)) {
      pd <- get_plot_data_inc(data, res[[m]])
      local_max <- max(c(pd$hist$y, pd$curve$y), na.rm = TRUE)
      if(local_max > max_y) max_y <- local_max
      all_plot_data[[m]] <- pd
    }
    max_y <- max_y * 1.05
    
    # sub-case a: overlaid (combined) view
    if (input$view_inc == "Overlaid") {
      pd <- all_plot_data[[1]] # base hist data
      
      hc <- highchart() %>%
        hc_chart(type = "line", style = list(fontFamily = "Tw Cen MT"), backgroundColor = "#FFFFFF") %>%
        hc_title(text = "Combined Model Comparison", style = list(fontSize="16px", fontWeight="bold")) %>%
        hc_plotOptions(
          series = list(pointPlacement = "on"), # fix for axis alignment
          column = list(grouping = FALSE, pointPadding = 0, groupPadding = 0, borderWidth = 0.25)
        ) %>%
        hc_xAxis(
          title = list(text = "Incubation Periods", style = axis_style), 
          allowDecimals = FALSE, 
          lineColor = "#000000", lineWidth = 1, tickWidth = 1, tickColor = "#000000",
          labels = list(style = label_style),
          min = 0, minPadding = 0, startOnTick = TRUE # fix for gap removal
        ) %>%
        hc_yAxis(
          title = list(text = "Density", style = axis_style), 
          gridLineWidth = 0, 
          lineWidth = 1, lineColor="#000000", tickWidth = 1, tickColor = "#000000",
          labels = list(style = label_style),
          min = 0, max = max_y, minPadding = 0, startOnTick = TRUE # fix for gap removal
        ) %>%
        # add histogram
        hc_add_series(
          data = list_parse(pd$hist),
          type = "column",
          name = "Observed Data",
          color = "#D3D3D3", 
          borderColor = "#FFFFFF", borderWidth = 0.25,
          pointRange = 1,
          zIndex = 0
        )
      
      # add curves
      for(m in names(res)) {
        nm <- tolower(m)
        pd_model <- all_plot_data[[m]]
        
        hc <- hc %>% hc_add_series(
          data = pd_model$curve,
          type = "line",
          name = paste(m, "Fit"),
          color = line_col_overlaid[[nm]], 
          lineWidth = 2, marker = list(enabled = FALSE),
          zIndex = 2
        )
      }
      
      hc <- hc %>% hc_exporting(enabled = TRUE, buttons = list(contextButton = list(menuItems = menu_items)))
      return(hc)
      
    } else {
      # sub-case b: side-by-side (composite) view
      hc <- highchart() %>%
        hc_chart(type = "column", style = list(fontFamily = "Tw Cen MT"), backgroundColor = "#FFFFFF") %>%
        hc_title(text = "Distribution Comparison", style = list(fontSize="16px", fontWeight="bold")) %>%
        hc_plotOptions(
          series = list(pointPlacement = "on"), # User Fix for alignment
          column = list(grouping = FALSE, pointPadding = 0, groupPadding = 0, borderWidth = 0.25)
        ) %>%
        
        # define 3 x-axes
        hc_xAxis_multiples(
          list(title = list(text = "Lognormal", style = axis_style), left = "0%", width = "30%", lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", labels=list(style = label_style), min=0, minPadding=0, startOnTick=TRUE),
          list(title = list(text = "Weibull", style = axis_style), left = "35%", width = "30%", offset = 0, lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", labels=list(style = label_style), min=0, minPadding=0, startOnTick=TRUE),
          list(title = list(text = "Gamma", style = axis_style), left = "70%", width = "30%", offset = 0, lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", labels=list(style = label_style), min=0, minPadding=0, startOnTick=TRUE)
        ) %>%
        
        # define 3 y-axes
        hc_yAxis_multiples(
          list(title = list(text = "Density", style = axis_style), left = "0%", width = "30%", lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", gridLineWidth=0, labels=list(style = label_style), min=0, max=max_y, minPadding=0, startOnTick=TRUE),
          list(title = list(text = "Density", style = axis_style), left = "35%", width = "30%", offset = 0, lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", gridLineWidth=0, labels=list(style = label_style), min=0, max=max_y, minPadding=0, startOnTick=TRUE),
          list(title = list(text = "Density", style = axis_style), left = "70%", width = "30%", offset = 0, lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", gridLineWidth=0, labels=list(style = label_style), min=0, max=max_y, minPadding=0, startOnTick=TRUE)
        ) %>%
        
        hc_exporting(
          enabled = TRUE,
          buttons = list(contextButton = list(menuItems = menu_items)),
          chartOptions = list(chart = list(backgroundColor = "#FFFFFF"))
        ) %>%
        hc_legend(enabled = FALSE)
      
      axes_idx <- list(lognormal=0, weibull=1, gamma=2)
      
      for(m in names(res)) {
        nm <- tolower(m)
        idx <- axes_idx[[nm]]
        if(is.null(idx)) next 
        
        pd <- all_plot_data[[m]]
        
        hc <- hc %>%
          hc_add_series(
            data = list_parse(pd$hist),
            type = "column",
            color = col_map[[nm]],
            xAxis = idx, yAxis = idx,
            borderColor = "#FFFFFF", borderWidth = 0.25,
            pointRange = 1, 
            name = paste(m, "Hist")
          ) %>%
          hc_add_series(
            data = pd$curve,
            type = "line",
            color = "#000000",
            xAxis = idx, yAxis = idx,
            lineWidth = 1.5, marker = list(enabled = FALSE),
            name = paste(m, "Fit")
          )
      }
      return(hc)
    }
    
  } else {
    # case 2: single model
    m <- names(res)[[1]]
    pd <- get_plot_data_inc(data, res[[1]])
    nm <- tolower(m)
    
    highchart() %>%
      hc_chart(style = list(fontFamily = "Tw Cen MT"), backgroundColor = "#FFFFFF") %>%
      hc_title(text = paste(str_to_title(m), "Distribution"), style = list(fontSize="16px", fontWeight="bold")) %>%      
      hc_plotOptions(
        series = list(pointPlacement = "on"), # fix for alignment
        column = list(grouping = FALSE, pointPadding = 0, groupPadding = 0, borderWidth = 0.25)
      ) %>%
      hc_add_series(
        data = list_parse(pd$hist),
        type = "column",
        name = "Histogram",
        color = col_map[[nm]], 
        borderColor = "#FFFFFF", borderWidth = 0.25,
        pointRange = 1
      ) %>%
      hc_add_series(
        data = pd$curve,
        type = "line",
        name = "Fit",
        color = "#000000",
        lineWidth = 1.5, marker = list(enabled = FALSE)
      ) %>%
      hc_xAxis(
        title = list(text = "Incubation Periods", style = axis_style), 
        min = floor(min(data))-1, tickInterval = 1,
        lineColor = "#000000", lineWidth = 1, tickWidth = 1, tickColor = "#000000",
        labels = list(style = label_style),
        min = 0, minPadding = 0, startOnTick = TRUE # fix for gap removal
      ) %>%
      hc_yAxis(
        title = list(text = "Density", style = axis_style),
        gridLineWidth = 0,
        lineWidth = 1, lineColor = "#000000", tickWidth = 1, tickColor = "#000000",
        labels = list(style = label_style),
        min = 0, minPadding = 0, startOnTick = TRUE # fix for gap removal
      ) %>%
      hc_exporting(
        enabled = TRUE,
        buttons = list(contextButton = list(menuItems = menu_items)),
        chartOptions = list(chart = list(backgroundColor = "#FFFFFF"))
      )
  }
})

```

### Dataset Preview

```{r inc-dataset-preview}
renderDataTable({
  dataset_inc()
}, options = list(
  pageLength = 10, 
  paging = TRUE, 
  striped = TRUE, 
  bordered = TRUE,
  hover = TRUE  
), rownames = FALSE)

```


Incubation Period (Bayesian) {data-orientation=rows}
==========================================

Sidebar {.sidebar}
--------------------

```{r inc-bayes-sidebar}

hr()
fileInput("file_bayes_inc", "Choose CSV File",
          multiple = FALSE,
          accept = c("text/csv", "text/comma-separated-values", ".csv"))

hr()
# stratification inputs
selectInput(inputId = "sex_bayes_inc", label = "Sex",
            choices = c("All", "Male", "Female"),
            selected = "All")

selectInput(inputId = "age_bayes_inc", label = "Age Group",
            choices = c("All", '0 - 19', '20 - 29', '30 - 39', '40 - 49', '50 - 59', '≥ 60'),
            selected = "All")

hr()
# model selection
selectInput("bayes_model_select", "Distribution Model", 
            choices = c("Lognormal","Weibull", "Gamma"), 
            selected = "Lognormal")

sliderInput("iter_count_inc", "Iterations (Max 4000)", value = 1000, min = 500, max = 4000, step = 100)

hr()
actionButton("run_bayes_inc", "Run Bayesian Model", icon = icon("play"), class = "btn-primary", width = "100%")

hr()
valueBox(value = tags$h6("Note: This model uses Bayesian MCMC (rstan). Calculations typically take 30-45 seconds.",
                         style = "text-align: justify; font-family: Tw Cen MT; font-size: 16px; color: white;"), 
         color = "#333333")

hr()

valueBox(
  value = downloadButton("downloadMockIncubationBayes", "Download Mock Data", icon = icon("download")),
  color = "#FFFFFF"
)

renderImage({
  pfad <- "one_logo.png"
  list(
    src = pfad,
    contentType = 'image/png',
    width = 210,
    height = 60,
    style = "position: fixed; bottom: 10px; left: 10px; cursor: pointer;",
    onclick = "window.location.href = '[http://onehealth.or.kr/](http://onehealth.or.kr/)';"
  )
}, deleteFile = F)

```


```{r inc-bayes-data-processing}

# global data loading (runs once per file upload)
bayes_inc_global <- reactive({
  req(input$file_bayes_inc)
  
  # robust read
  tryCatch({
    df <- read.csv(input$file_bayes_inc$datapath, stringsAsFactors = FALSE)
  }, error = function(e) {
    validate(need(FALSE, "error reading file. please ensure it is a valid csv."))
  })
  
  df <- clean_names(df) 
  
  # validation start
  validate(
    need(all(c("symptom_onset", "exp_date_min", "exp_date_max") %in% names(df)),
         "incorrect dataset: please upload incubation bayesian data containing 'symptom onset', 'exp date min', and 'exp date max'.")
  )
  # validation end
  
  # standardize sex
  if("sex" %in% names(df)) {
    df <- df %>% mutate(sex = case_when(
        grepl("^m", sex, ignore.case = TRUE) ~ "Male",
        grepl("^f", sex, ignore.case = TRUE) ~ "Female",
        TRUE ~ sex
      ))
  }
  
  # standardize age group
  if("age" %in% names(df)) {
    df$age <- suppressWarnings(as.numeric(df$age))
    df <- df %>% mutate(age_group = case_when(
      age > 0 & age <= 19 ~ '0 - 19',
      age >= 20 & age <= 29 ~ '20 - 29',
      age >= 30 & age <= 39 ~ '30 - 39',
      age >= 40 & age <= 49 ~ '40 - 49',
      age >= 50 & age <= 59 ~ '50 - 59',
      TRUE ~ '≥ 60'
    ))
  }
  
  # parse dates globally
  df <- df %>%
    mutate(
      date_symp = parse_date(symptom_onset),
      date_min  = parse_date(exp_date_min),
      date_max  = parse_date(exp_date_max)
    )
  
  df
})

# filtering logic
bayes_inc_data <- reactive({
  req(bayes_inc_global())
  df <- bayes_inc_global()
  
  # filter raw data
  if (input$sex_bayes_inc != "All" && "sex" %in% names(df)) {
    df <- df %>% filter(sex == input$sex_bayes_inc)
  }
  if (input$age_bayes_inc != "All" && "age_group" %in% names(df)) {
    df <- df %>% filter(age_group == input$age_bayes_inc)
  }
  
  # capture count before cleaning
  n_before_cleaning <- nrow(df)
  
  # model prep logic
  df <- df %>%
    filter(!is.na(date_symp)) %>% # must have symptom date
    mutate(
      tSymptomOnset = as.integer(date_symp - as.Date("2019-12-31")),
      tStartExposure = as.integer(date_min - as.Date("2019-12-31")),
      tEndExposure = as.integer(date_max - as.Date("2019-12-31"))
    ) %>%
    # logic update!!
    # keep rows with at least one exposure date
    filter(!is.na(tStartExposure) | !is.na(tEndExposure)) %>%
    # impute missing end date (assume exposure continued until symptoms)
    mutate(tEndExposure = ifelse(is.na(tEndExposure), tSymptomOnset, tEndExposure)) %>%
    # impute missing start date (assume exposure started 14 days before end date)
    mutate(tStartExposure = ifelse(is.na(tStartExposure), tEndExposure - 14, tStartExposure)) %>% # you can update the window here
    # final safety check: end exposure cannot be after symptoms
    mutate(tEndExposure = ifelse(tEndExposure > tSymptomOnset, tSymptomOnset, tEndExposure)) %>%
    # ensure start is before end
    filter(tStartExposure <= tEndExposure)
  
  # capture count after
  n_final <- nrow(df)
  n_dropped <- n_before_cleaning - n_final
  
  # notify user
  if (n_dropped > 0) {
    showNotification(
      paste0("data loaded: ", n_final, " valid rows used. (", n_dropped, " rows removed: missing both exposure dates or invalid timeline)"),
      type = "warning",
      duration = 5 
    )
  } else {
    showNotification(
      paste0("data successfully loaded: ", n_final, " rows ready."),
      type = "message",
      duration = 3 
    )
  }
  
  df
})

# modeling
bayes_inc_fit <- eventReactive(input$run_bayes_inc, {
  
  # lock button
  shinyjs::disable("run_bayes_inc")
  
  # unlock button when done
  on.exit({
    shinyjs::enable("run_bayes_inc")
  })
  
  # double clean
  gc()
  gc() 
  
  req(bayes_inc_data())
  df <- bayes_inc_data()

  validate(
    need(nrow(df) >= 5, "insufficient data: you need at least 5 observations to run this model.")
  )

  stan_data <- list(
    N = nrow(df),
    tStartExposure = df$tStartExposure,
    tEndExposure = df$tEndExposure,
    tSymptomOnset = df$tSymptomOnset
  )

  withProgress(message = 'running bayesian model...', detail = 'initializing...', value = 0, {

    mod <- get_compiled_stan_model(input$bayes_model_select)
    incProgress(0.1, detail = "sampling chains...")

    fit <- rstan::sampling(
      object = mod,
      data = stan_data,
      chains = 1,
      cores = 1,
      iter = input$iter_count_inc,
      warmup = input$iter_count_inc / 2,
      refresh = 0,
      pars = c("alphaInc", "sigmaInc"), # keep light!
      control = list(adapt_delta = 0.95)
    )

    incProgress(0.8, detail = "extracting posteriors...")

    params <- rstan::extract(fit)

    list(
      fit_obj = fit,
      iterations = input$iter_count_inc,
      model_name = input$bayes_model_select,
      alpha = params$alphaInc,
      sigma = params$sigmaInc,
      alpha_summary = quantile(params$alphaInc, probs = c(0.025, 0.5, 0.975)),
      sigma_summary = quantile(params$sigmaInc, probs = c(0.025, 0.5, 0.975))
    )
  })
})

```

## Row {data-height="180"}

### Sample Size

```{r inc-bayes-sample-size}

renderUI({
  req(bayes_inc_data())
  
  ss <- nrow(bayes_inc_data())
  
  htmltools::div(
    style = "display: flex; flex-direction: column; justify-content: center; align-items: center; height: 100%; width: 100%; padding: 10px;",
    htmltools::div(
      style = "font-size: 40px; font-weight: bold; margin-bottom: 5px; text-align: center; font-family: Tw Cen MT; color: #000000;",
      ss
    ),
    htmltools::div(
      style = "font-size: 14px; text-align: center; color: #666;",
      "Observations"
    )
  )
})

```


### Model Statistics (Median & 95% CrI)

```{r inc-bayes-model-stats}

renderDT({
  req(bayes_inc_fit())
  res <- bayes_inc_fit()
  
  # dynamic parameter labels & mean calculation
  if (res$model_name == "Lognormal") {
    p_names <- c("MeanLog (Alpha)", "SDLog (Sigma)")
    # mean of lognormal = exp(mu + sigma^2/2)
    post_means <- exp(res$alpha + (res$sigma^2)/2)
  } else if (res$model_name == "Gamma") {
    p_names <- c("Shape (Alpha)", "Rate (Sigma)")
    # mean of gamma = shape / rate
    post_means <- res$alpha / res$sigma
  } else {
    # weibull
    p_names <- c("Shape (Alpha)", "Scale (Sigma)")
    # mean of weibull = scale * gamma(1 + 1/shape)
    post_means <- res$sigma * gamma(1 + 1/res$alpha)
  }
  
  df_stats <- data.frame(
    Parameter = p_names,
    `Median` = round(c(res$alpha_summary[2], res$sigma_summary[2]), 3),
    `Lower_95_CrI` = round(c(res$alpha_summary[1], res$sigma_summary[1]), 3),
    `Upper_95_CrI` = round(c(res$alpha_summary[3], res$sigma_summary[3]), 3)
  )
  
  mean_summ <- quantile(post_means, probs = c(0.025, 0.5, 0.975))
  
  df_mean <- data.frame(
    Parameter = "Mean Incubation Period",
    `Median` = round(mean_summ[2], 3),
    `Lower_95_CrI` = round(mean_summ[1], 3),
    `Upper_95_CrI` = round(mean_summ[3], 3)
  )
  
  final_df <- rbind(df_stats, df_mean)
  
  fname <- paste0("Incubation_Bayesian_Stats_", Sys.Date())
  
  datatable(
    final_df, 
    rownames = FALSE, 
    extensions = 'Buttons',
    options = list(
      dom = 'Brt',
      buttons = list(
        list(extend = "csv", filename = fname),
        list(extend = "excel", filename = fname),
        list(extend = "pdf", filename = fname)
      ),
      columnDefs = list(list(className = 'dt-center', targets = "_all"))
    )
  ) %>%
    formatStyle(columns = "Parameter", fontWeight = "bold")
})

```

Column {.tabset .tabset-fade }
-------------------------------


### Dataset Preview

```{r inc-bayes-dataset-preview}

renderDataTable({
  bayes_inc_data()
}, options = list(
  pageLength = 10, 
  paging = TRUE, 
  striped = TRUE, 
  bordered = TRUE, 
  hover = TRUE  
), rownames = FALSE)

```


### Posterior Distribution (Median Fit)

```{r inc-bayes-posterior-dist}

renderHighchart({
  req(bayes_inc_fit())
  res <- bayes_inc_fit()
  
  alpha_est <- median(res$alpha)
  sigma_est <- median(res$sigma)
  x_vals <- seq(0, 30, length.out = 200)
  
  # dynamic distribution logic
  if (res$model_name == "Gamma") {
    y_vals <- dgamma(x_vals, shape = alpha_est, rate = sigma_est)
  } else if (res$model_name == "Lognormal") {
    y_vals <- dlnorm(x_vals, meanlog = alpha_est, sdlog = sigma_est)
  } else {
    y_vals <- dweibull(x_vals, shape = alpha_est, scale = sigma_est)
  }
  
  model_label <- tools::toTitleCase(res$model_name)
  menu_items <- list("viewFullscreen", "separator", "downloadPNG", "downloadPDF", "downloadSVG")
  
  # theme consistency
  axis_title_style <- list(color = "#000000", fontSize = "16px", fontWeight = "bold", fontFamily = "Tw Cen MT")
  axis_label_style <- list(color = "#000000", fontSize = "14px", fontWeight = "bold", fontFamily = "Tw Cen MT")
  
  highchart() %>%
    hc_chart(style = list(fontFamily = "Tw Cen MT"), backgroundColor = "#FFFFFF") %>%
    
    hc_title(
      text = paste0("Posterior Predictive Distribution (", model_label, ")"),
      style = list(fontSize = "18px", fontWeight = "bold", fontFamily = "Tw Cen MT", color = "#000000")
    ) %>%
    
    hc_subtitle(
      text = paste0("Based on ", res$iterations, " MCMC iterations"),
      style = list(fontSize = "14px", fontFamily = "Tw Cen MT", color = "#666666")
    ) %>%
    
    hc_exporting(
      enabled = TRUE, 
      filename = paste0("Incubation_Posterior_Distribution_", Sys.Date()),
      buttons = list(contextButton = list(menuItems = menu_items))
    ) %>%
    
    hc_add_series(
      data = data.frame(x = x_vals, y = y_vals),
      type = "area",
      name = "Posterior Density",
      color = "rgba(65, 105, 225, 0.5)",
      marker = list(enabled = FALSE),      
      lineWidth = 2
    ) %>%
    
    # axis styling
    hc_xAxis(
      title = list(text = "Incubation Period (Days)", style = axis_title_style),
      labels = list(style = axis_label_style),
      min = 0, minPadding = 0, startOnTick = TRUE,
      lineColor = "#000000", lineWidth = 1, tickColor = "#000000"
    ) %>%
    
    hc_yAxis(
      title = list(text = "Density", style = axis_title_style),
      labels = list(style = axis_label_style),
      min = 0, minPadding = 0, startOnTick = TRUE,
      gridLineWidth = 0, 
      lineColor = "#000000", lineWidth = 1, tickColor = "#000000"
    ) %>%
    
    hc_tooltip(valueDecimals = 4, style = list(fontFamily = "Tw Cen MT"))
})

```

### Trace Plot (Convergence Check)

```{r inc-bayes-trace-plot}

renderPlotly({
  req(bayes_inc_fit())
  
  # base plot with grid lines but customized fonts and black axis
  p <- rstan::traceplot(bayes_inc_fit()$fit_obj, 
                        pars = c("alphaInc", "sigmaInc"),
                        linewidth = 0.25) + 
    theme_bw() + 
    ggtitle("Trace Plot: Alpha & Sigma") +
    theme(
      text = element_text(family = "Tw Cen MT"),
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      axis.title = element_text(size = 12, face = "bold"),
      axis.text = element_text(size = 10, color = "black"),
      axis.line = element_line(color = "black"), 
      strip.background = element_rect(fill = "#f0f0f0"), 
      strip.text = element_text(face = "bold")
    )
  
  ggplotly(p) %>%
    config(
      toImageButtonOptions = list(
        format = "png",
        filename = paste0("Incubation_Trace_Plot_", Sys.Date()),
        height = 500,
        width = 700,
        scale = 2 
      )
    ) %>%
    layout(font = list(family = "Tw Cen MT"))
})

```

Serial Interval {data-orientation=rows}
=========================================

Sidebar {.sidebar}
---------------------

```{r si-sidebar}
hr()
fileInput("file1_si", "Choose CSV File",
          multiple = TRUE,
          accept = c("text/csv", "text/comma-separated-values,text/plain", ".csv"))
hr()
selectInput(inputId = "sex_si", label = "Sex",
            choices = c("All", "Male", "Female"),
            selected = "All")

selectInput(inputId = "age_group_si", label = "Age Group",
            choices = c("All", '0 - 19', '20 - 29', '30 - 39', '40 - 49', '50 - 59', '≥ 60'),
            selected = "All")
hr()

selectInput("model_si", "Distribution Model:",
            choices = c("Normal", "Weibull", "Gamma", "Lognormal", "ALL"),
            selected = "Normal")

# comparison view toggle for si
conditionalPanel(
  condition = "input.model_si == 'ALL'",
  radioButtons("view_si", "Comparison View:",
               choices = c("Side-by-Side", "Overlaid"),
               selected = "Side-by-Side")
)

hr()

valueBox(
  value = downloadButton("downloadMockserial", "Download Mock Data", icon = icon("download")),
  color = "#FFFFFF"
)

hr()

valueBox(
  value = tags$h6("Note: Parametric models (Weibull/Gamma/Lognormal) are fit on (Data + 1) to handle zeros, then estimates are shifted back. Normal model is fit on raw data.",
                  style = "text-align: justify; font-family: Tw Cen MT; font-size: 16px; color: white;"),
  color = "#FFFFFF")

hr()
# logo image
renderImage({
  list(
    src = "one_logo.png",
    contentType = 'image/png',
    width = 210,
    height = 60,
    style = "position: fixed; bottom: 10px; left: 10px; cursor: pointer;",
    onclick = "window.location.href = '[http://onehealth.or.kr/](http://onehealth.or.kr/)';"
  )
}, deleteFile = FALSE)

```


```{r si-data-processing}

# data processing for serial interval
dataset_si <- eventReactive(list(input$file1_si, input$sex_si, input$age_group_si), {
  req(input$file1_si)
  
  dataset <- read.csv(input$file1_si$datapath)
  dataset <- data.frame(lapply(dataset, trimws), stringsAsFactors = FALSE)
  dataset <- clean_names(dataset)
  
  # validation start
  validate(
    need(all(c("id", "source_of_infection", "date_of_onset") %in% names(dataset)),
         "incorrect dataset: please upload serial interval data containing 'id', 'source of infection', and 'date of onset'.")
  )
  # validation end
  
  # more robust cleaning
  if("age_group" %in% names(dataset)) {
    dataset$age_group <- dataset$age_group %>%
      gsub("s$", "", ., ignore.case = TRUE) %>%
      gsub("[^0-9]", "", .)
    dataset$age_group <- suppressWarnings(as.numeric(dataset$age_group))
    dataset <- dataset %>% mutate(age_group = case_when(
      age_group > 0 & age_group <= 19 ~ '0 - 19',
      age_group >= 20 & age_group <= 29 ~ '20 - 29',
      age_group >= 30 & age_group <= 39 ~ '30 - 39',
      age_group >= 40 & age_group <= 49 ~ '40 - 49',
      age_group >= 50 & age_group <= 59 ~ '50 - 59',
      TRUE ~ '≥ 60'
    ))
  }
  
  if("sex" %in% names(dataset)) {
    dataset <- dataset %>%
      mutate(sex = case_when(
        grepl("^m", sex, ignore.case = TRUE) ~ "Male",
        grepl("^f", sex, ignore.case = TRUE) ~ "Female",
        TRUE ~ sex
      ))
  }
  
  dataset$id <- as.numeric(gsub("[^0-9]", "", dataset$id))
  dataset$source_of_infection <- as.numeric(gsub("[^0-9]", "", dataset$source_of_infection))
  dataset$date_of_exposure <- parse_date(dataset$date_of_exposure)
  dataset$date_of_onset <- parse_date(dataset$date_of_onset)
  dataset$source_date <- dataset$date_of_onset[match(dataset$source_of_infection, dataset$id)]
  
  if (input$sex_si != "All") dataset <- dataset %>% filter(sex == input$sex_si)
  if (input$age_group_si != "All") dataset <- dataset %>% filter(age_group == input$age_group_si)
  
  dataset
})

get_serial_data <- function() {
  req(dataset_si())
  df <- dataset_si()
  
  required_cols <- c("id", "source_of_infection", "date_of_onset", "source_date")
  if (!all(required_cols %in% names(df))) stop("required columns not found in dataset")
  
  df <- df %>%
    filter(!is.na(date_of_onset), !is.na(source_date)) %>%
    mutate(serial = as.numeric(date_of_onset - source_date)) %>%
    filter(!is.na(serial))
  
  if (nrow(df) == 0) return(NULL)
  
  df$serial
}

# master fitting logic for si
fit_results_si <- reactive({
  data <- get_serial_data()
  if (is.null(data) || length(data) < 3) return(NULL)
  
  models <- if(input$model_si == "ALL") c("normal", "weibull", "gamma", "lognormal") else tolower(input$model_si)
  
  res <- list()
  for(m in models) {
    # shift logic: normal model=0, others=1
    shift_val <- if(m == "normal") 0 else 1
    
    # for non-normal models, we must ensure data + shift > 0.
    if(m != "normal") {
      valid_indices <- (data + shift_val) > 0
      data_for_fit <- data[valid_indices]
    } else {
      data_for_fit <- data
    }
    
    # check if we still have enough data after filtering
    if(length(data_for_fit) < 3) next
    
    suppressWarnings({
      fit <- fit_parametric_model(data_for_fit, m, shift_val = shift_val)
    })
    
    if(!is.null(fit)) res[[m]] <- fit
  }
  res
})

```


## Row {data-height="180"}

### Sample Size

```{r si-sample-size}

renderUI({
  req(dataset_si())
  ss <- nrow(dataset_si())
  htmltools::div(
    style = "display: flex; flex-direction: column; justify-content: center; align-items: center; height: 100%; width: 100%; padding: 10px;",
    htmltools::div(
      style = "font-size: 30px; font-weight: bold; margin-bottom: 10px; text-align: center;font-family: Tw Cen MT;",
      ss
    )
  )
})

```

### Model Fit Statistics

```{r si-model-fit-stats}

renderDT({
  res <- fit_results_si()
  data <- get_serial_data()
  
  if (!is.null(data) && length(data) < 3) {
    df <- data.frame(Status = "Insufficient Data (N < 3).")
  } else if (is.null(res) || length(res) == 0) {
    df <- data.frame(Status = "No data or fit failed")
  } else {
    if (input$model_si != "ALL") {
      df <- res[[1]]$raw_params_df
    } else {
      df <- do.call(rbind, lapply(res, function(x) {
        data.frame(
          Model = x$model, 
          AIC = round(x$aic, 2), 
          BIC = round(x$bic, 2)
        )
      }))
      df <- df[order(df$AIC), ]
    }
  }
  
  datatable(
    df, rownames = FALSE,
    options = list(dom = "Brt", buttons = list("csv", "excel", "pdf"), pageLength = nrow(df)),
    extensions = "Buttons"
  ) %>% 
    formatStyle(columns = names(df), `font-size` = "12px") %>%
    formatStyle(columns = intersect(names(df), c("parameter", "Model")), fontWeight = "bold")
})

```

### Model Parameters

```{r si-model-params}

renderDT({
  res <- fit_results_si()
  data <- get_serial_data()
  
  if (!is.null(data) && length(data) < 3) {
    df <- data.frame(Status = "Insufficient Data")
  } else if (is.null(res) || length(res) == 0) {
    df <- data.frame(Status = "No data")
  } else {
    fmt_moments <- function(x) {
      data.frame(
        parameter = c(paste(x$model, "Mean"), paste(x$model, "Variance"), "95th Percentile"),
        value = round(c(x$moments["mean"], x$moments["var"], x$moments["ci95"]), 3)
      )
    }
    
    if (input$model_si != "ALL") {
      df <- fmt_moments(res[[1]])
    } else {
      df <- do.call(rbind, lapply(res, function(x) {
        data.frame(
          Model = x$model,
          `Mean` = round(x$moments["mean"], 3),
          `Variance` = round(x$moments["var"], 3),
          `95th Percentile` = round(x$moments["ci95"], 3),
          check.names = FALSE
        )
      }))
    }
  }
  
  datatable(
    df, rownames = FALSE,
    options = list(dom = "Brt", buttons = list("csv", "excel", "pdf"), pageLength = nrow(df)),
    extensions = "Buttons"
  ) %>% 
    formatStyle(columns = names(df), `font-size` = "12px") %>%
    formatStyle(columns = intersect(names(df), c("parameter", "Model")), fontWeight = "bold")
})

```

Column {.tabset .tabset-fade }
-------------------------------

### Distribution Plots

```{r si-histograms}

renderHighchart({
  data <- get_serial_data()
  res <- fit_results_si()
  
  if (is.null(data) || length(data) < 3 || is.null(res) || length(res) == 0) {
    return(highchart() %>% hc_title(text="Insufficient Data (N < 3)"))
  }
  
  # set model colors
  col_map <- list(
    normal = "#DAA520",    # gold
    lognormal = "#00BFFF", # blue
    weibull = "#FF8C00",   # orange
    gamma = "#228B22"      # green
  )
  line_col_overlaid <- list(normal="#0000FF", lognormal="#000000", weibull="#FF0000", gamma="#008000")
  
  # style definitions
  axis_style <- list(color = "#000000", fontSize = "16px", fontWeight = "bold")
  label_style <- list(color = "#000000", fontSize = "14px", fontWeight = "bold")
  menu_items <- list("viewFullscreen", "separator", "downloadPNG", "downloadPDF", "downloadSVG")
  
  # ensure the first bar isn't cut in half (start x-axis slightly before min integer)
  min_x_val <- floor(min(data)) - 0.5
  
  # case 1: all models selected
  if (input$model_si == "ALL") {
    
    # calculate global max y
    max_y <- 0
    all_plot_data <- list()
    for(m in names(res)) {
      pd <- get_plot_data_si(data, res[[m]])
      local_max <- max(c(pd$hist$y, pd$curve$y), na.rm = TRUE)
      if(local_max > max_y) max_y <- local_max
      all_plot_data[[m]] <- pd
    }
    max_y <- max_y * 1.05
    
    # sub-case a: overlaid (combined) view
    if (input$view_si == "Overlaid") {
      
      pd <- all_plot_data[[1]] # base hist
      
      hc <- highchart() %>%
        hc_chart(type = "line", style = list(fontFamily = "Tw Cen MT"), backgroundColor = "#FFFFFF") %>%
        hc_title(text = "Combined Model Comparison", style = list(fontSize="16px", fontWeight="bold")) %>%
        hc_plotOptions(
          series = list(pointPlacement = "on"), 
          column = list(grouping = FALSE, pointPadding = 0, groupPadding = 0, borderWidth = 0.25)
        ) %>%
        hc_xAxis(
          title = list(text = "Serial Interval (Days)", style = axis_style), 
          allowDecimals = FALSE, 
          lineColor = "#000000", lineWidth = 1, tickWidth = 1, tickColor = "#000000",
          labels = list(style = label_style),
          min = min_x_val, # Updated start point
          minPadding = 0, startOnTick = TRUE
        ) %>%
        hc_yAxis(
          title = list(text = "Density", style = axis_style), 
          gridLineWidth = 0, 
          lineWidth = 1, lineColor="#000000", tickWidth = 1, tickColor = "#000000",
          labels = list(style = label_style),
          min = 0, max = max_y, 
          minPadding = 0, startOnTick = TRUE
        ) %>%
        hc_add_series(
          data = list_parse(pd$hist), type = "column", name = "Observed Data",
          color = "#D3D3D3", borderColor = "#FFFFFF", borderWidth = 0.25,
          pointRange = 2, zIndex = 0
        )
      
      for(m in names(res)) {
        nm <- tolower(m)
        pd_model <- all_plot_data[[m]]
        hc <- hc %>% hc_add_series(
          data = pd_model$curve, type = "line", name = paste(m, "Fit"),
          color = line_col_overlaid[[nm]], lineWidth = 2, marker = list(enabled = FALSE), zIndex = 2
        )
      }
      hc <- hc %>% hc_exporting(enabled = TRUE, buttons = list(contextButton = list(menuItems = menu_items)))
      return(hc)
      
    } else {
      # sub-case b: side-by-side (composite) view with constant scale
      hc <- highchart() %>%
        hc_chart(type = "column", style = list(fontFamily = "Tw Cen MT"), backgroundColor = "#FFFFFF") %>%
        hc_title(text = "Distribution Comparison", style = list(fontSize="16px", fontWeight="bold")) %>%
        hc_plotOptions(
          series = list(pointPlacement = "on"),
          column = list(grouping = FALSE, pointPadding = 0, groupPadding = 0, borderWidth = 0.25)
        ) %>%
        
        # 4 x-axes
        hc_xAxis_multiples(
          list(title = list(text = "Normal", style=axis_style), left = "0%", width = "22%", lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", labels=list(style=label_style), min=min_x_val, minPadding=0, startOnTick=TRUE),
          list(title = list(text = "Weibull", style=axis_style), left = "26%", width = "22%", offset = 0, lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", labels=list(style=label_style), min=min_x_val, minPadding=0, startOnTick=TRUE),
          list(title = list(text = "Gamma", style=axis_style), left = "52%", width = "22%", offset = 0, lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", labels=list(style=label_style), min=min_x_val, minPadding=0, startOnTick=TRUE),
          list(title = list(text = "Lognormal", style=axis_style), left = "78%", width = "22%", offset = 0, lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", labels=list(style=label_style), min=min_x_val, minPadding=0, startOnTick=TRUE)
        ) %>%
        
        # 4 y-axes (aligned, same scale)
        hc_yAxis_multiples(
          list(title = list(text = "Density", style=axis_style), left = "0%", width = "22%", lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", gridLineWidth=0, labels=list(style=label_style), min=0, max=max_y, minPadding=0, startOnTick=TRUE),
          list(title = list(text = "Density", style=axis_style), left = "26%", width = "22%", offset = 0, lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", gridLineWidth=0, labels=list(style=label_style), min=0, max=max_y, minPadding=0, startOnTick=TRUE),
          list(title = list(text = "Density", style=axis_style), left = "52%", width = "22%", offset = 0, lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", gridLineWidth=0, labels=list(style=label_style), min=0, max=max_y, minPadding=0, startOnTick=TRUE),
          list(title = list(text = "Density", style=axis_style), left = "78%", width = "22%", offset = 0, lineWidth=1, lineColor="black", tickWidth=1, tickColor="black", gridLineWidth=0, labels=list(style=label_style), min=0, max=max_y, minPadding=0, startOnTick=TRUE)
        ) %>%
        
        hc_exporting(enabled = TRUE, buttons = list(contextButton = list(menuItems = menu_items)), chartOptions = list(chart = list(backgroundColor = "#FFFFFF"))) %>%
        hc_legend(enabled = FALSE)
      
      axes_idx <- list(normal=0, weibull=1, gamma=2, lognormal=3)
      for(m in names(res)) {
        nm <- tolower(m)
        idx <- axes_idx[[nm]]
        if(is.null(idx)) next 
        
        pd <- all_plot_data[[m]]
        
        hc <- hc %>%
          hc_add_series(
            data = list_parse(pd$hist), type = "column", color = col_map[[nm]],
            xAxis = idx, yAxis = idx, 
            borderColor = "#FFFFFF", borderWidth = 0.25,
            pointRange = 2, 
            name = paste(m, "Hist")
          ) %>%
          hc_add_series(
            data = pd$curve, type = "line", color = "#000000",
            xAxis = idx, yAxis = idx, 
            lineWidth = 1.5, marker = list(enabled = FALSE),
            name = paste(m, "Fit")
          )
      }
      return(hc)
    }
    
  } else {
    # case 2: single model plot
    m <- names(res)[[1]]
    pd <- get_plot_data_si(data, res[[1]])
    nm <- tolower(m)
    
    highchart() %>%
      hc_chart(style = list(fontFamily = "Tw Cen MT"), backgroundColor = "#FFFFFF") %>%
      hc_title(text = paste(str_to_title(m), "Distribution"), style = list(fontSize="16px", fontWeight="bold")) %>%      
      hc_plotOptions(
        series = list(pointPlacement = "on"), 
        column = list(grouping = FALSE, pointPadding = 0, groupPadding = 0, borderWidth = 0.25)
      ) %>%
      hc_add_series(
        data = list_parse(pd$hist), type = "column", name = "Histogram",
        color = col_map[[nm]], borderColor = "#FFFFFF", borderWidth = 0.25,
        pointRange = 2
      ) %>%
      hc_add_series(
        data = pd$curve, type = "line", name = "Fit",
        color = "#000000", lineWidth = 1.5, marker = list(enabled = FALSE)
      ) %>%
      hc_xAxis(
        title = list(text = "Serial interval (days)", style = axis_style), 
        min = min_x_val, tickInterval = 2,
        lineColor = "#000000", lineWidth = 1, tickWidth = 1, tickColor = "#000000",
        labels = list(style = label_style),
        minPadding = 0, startOnTick = TRUE
      ) %>%
      hc_yAxis(
        title = list(text = "Density", style = axis_style),
        gridLineWidth = 0, lineWidth = 1, lineColor = "#000000", tickWidth = 1, tickColor = "#000000",
        labels = list(style = label_style),
        min = 0, minPadding = 0, startOnTick = TRUE
      ) %>%
      hc_exporting(enabled = TRUE, buttons = list(contextButton = list(menuItems = menu_items)), chartOptions = list(chart = list(backgroundColor = "#FFFFFF")))
  }
})

```

### Dataset Preview

```{r}

renderDataTable({
  req(dataset_si())
  dataset <- dataset_si()
  
  # ensure date columns are formatted as strings in yyyy-mm-dd
  dataset$date_of_exposure <- as.Date(dataset$date_of_exposure)
  dataset$date_of_onset <- as.Date(dataset$date_of_onset)
  dataset$source_date <- as.Date(dataset$source_date)
  
  datatable(
    dataset,
    options = list(
      pageLength = 10,
      paging = TRUE,
      striped = TRUE,
      bordered = TRUE,
      hover = TRUE
    ),
    rownames = FALSE
  )
})

```

Serial Interval (Bayesian) {data-orientation=rows}
==========================================

Sidebar {.sidebar}
--------------------

```{r si-bayes-sidebar}

hr()
fileInput("file_bayes_si", "Choose CSV File",
          multiple = FALSE,
          accept = c("text/csv", "text/comma-separated-values", ".csv"))

hr()
# stratification inputs
selectInput(inputId = "sex_bayes_si", label = "Sex",
            choices = c("All", "Male", "Female"),
            selected = "All")

selectInput(inputId = "age_bayes_si", label = "Age Group",
            choices = c("All", '0 - 19', '20 - 29', '30 - 39', '40 - 49', '50 - 59', '≥ 60'),
            selected = "All")

hr()
# model selection
selectInput("bayes_model_select_si", "Distribution Model", 
            choices = c("Normal","Weibull", "Gamma", "Lognormal"), 
            selected = "Normal")

sliderInput("iter_count_si", "Iterations (Max 4000)", value = 1000, min = 500, max = 4000, step = 100)

hr()
actionButton("run_bayes_si", "Run Bayesian Model", icon = icon("play"), class = "btn-primary", width = "100%")

hr()
valueBox(value = tags$h6("Note: This model estimates Serial Interval by linking Infectors to Infectees. Ensure 'Source of infection' IDs match 'id'.",
                         style = "text-align: justify; font-family: Tw Cen MT; font-size: 16px; color: white;"), 
         color = "#333333")

hr()

valueBox(
  value = downloadButton("downloadMockSerialBayes", "Download Mock Data", icon = icon("download")),
  color = "#FFFFFF"
)

renderImage({
  pfad <- "one_logo.png"
  list(
    src = pfad,
    contentType = 'image/png',
    width = 210,
    height = 60,
    style = "position: fixed; bottom: 10px; left: 10px; cursor: pointer;",
    onclick = "window.location.href = '[http://onehealth.or.kr/](http://onehealth.or.kr/)';"
  )
}, deleteFile = F)

```


```{r si-bayes-data-processing}

# data loading/processing
bayes_si_global <- reactive({
  req(input$file_bayes_si)
  tryCatch({
    df <- read.csv(input$file_bayes_si$datapath, stringsAsFactors = FALSE)
  }, error = function(e) {
    validate(need(FALSE, "error reading file. please ensure it is a valid csv."))
  })
  names(df) <- tolower(names(df))
  names(df) <- gsub("\\.", "_", names(df))
  
  # validation start
  validate(
    need(all(c("id", "source_of_infection", "date_of_onsetmin", "date_of_onsetmax") %in% names(df)),
         "incorrect dataset: please upload serial bayesian data containing 'id', 'source of infection', 'date of onsetmin', and 'date of onsetmax'.")
  )
  # validation end
  
  # standardize sex
  if("sex" %in% names(df)) {
    df <- df %>% mutate(sex = case_when(
      grepl("^m", sex, ignore.case = TRUE) ~ "Male",
      grepl("^f", sex, ignore.case = TRUE) ~ "Female",
      TRUE ~ sex
    ))
  }
  
  # standardize age group
  col_age <- grep("age.*group", names(df), value = TRUE)[1]
  if(!is.na(col_age)) {
    df$age_clean <- df[[col_age]]
    df <- df %>% mutate(age_group = case_when(
      grepl("0-19|under 10|10s", age_clean, ignore.case=T) ~ '0 - 19',
      grepl("20s|20-29", age_clean, ignore.case=T) ~ '20 - 29',
      grepl("30s|30-39", age_clean, ignore.case=T) ~ '30 - 39',
      grepl("40s|40-49", age_clean, ignore.case=T) ~ '40 - 49',
      grepl("50s|50-59", age_clean, ignore.case=T) ~ '50 - 59',
      grepl("60s|70s|80s|60", age_clean, ignore.case=T) ~ '≥ 60',
      TRUE ~ 'Unknown'
    ))
  }
  
  df$id_clean <- trimws(gsub("#", "", df$id))
  df$source_clean <- trimws(gsub("#", "", df$source_of_infection))
  df$onset_min <- parse_date(df$date_of_onsetmin)
  df$onset_max <- parse_date(df$date_of_onsetmax)
  
  df
})

# filtering logic 
bayes_si_data <- reactive({
  req(bayes_si_global())
  
  # clean memory
  gc() 
  
  df <- bayes_si_global()
  
  df_infectors <- df 
  df_infectees <- df
  
  if (input$sex_bayes_si != "All" && "sex" %in% names(df)) {
    df_infectees <- df_infectees %>% filter(sex == input$sex_bayes_si)
  }
  if (input$age_bayes_si != "All" && "age_group" %in% names(df)) {
    df_infectees <- df_infectees %>% filter(age_group == input$age_bayes_si)
  }
  
  n_before_cleaning <- nrow(df_infectees)
  
  df_merged <- df_infectees %>%
    inner_join(df_infectors %>% dplyr::select(id_clean, onset_min, onset_max), 
               by = c("source_clean" = "id_clean"), 
               suffix = c("_infectee", "_infector"))
  
  df_final <- df_merged %>%
    filter(!is.na(onset_min_infectee), !is.na(onset_min_infector)) %>%
    mutate(
      tStartExposure = as.integer(onset_min_infector - as.Date("2019-12-31")),
      tEndExposure = as.integer(onset_max_infector - as.Date("2019-12-31")),
      tSymptomOnset = as.integer(onset_min_infectee - as.Date("2019-12-31")) + 
        as.integer(onset_max_infectee - onset_min_infectee)/2
    ) 
  
  if (input$bayes_model_select_si != "Normal") {
    df_final <- df_final %>% filter(tSymptomOnset > tStartExposure)
  }
  
  n_final <- nrow(df_final)
  n_dropped <- n_before_cleaning - n_final
  
  if (n_dropped > 0) {
    showNotification(
      paste0("data loaded: ", n_final, " linked pairs found. (", n_dropped, " rows removed: unlinked or invalid dates)"),
      type = "warning", 
      duration = 5
    )
  } else {
    showNotification(
      paste0("data successfully loaded: ", n_final, " pairs ready."),
      type = "message", 
      duration = 3
    )
  }
  
  df_final
})

# modeling
bayes_si_fit <- eventReactive(input$run_bayes_si, {
  
  shinyjs::disable("run_bayes_si")
  
  on.exit({
    shinyjs::enable("run_bayes_si")
  })
  
  # cleanup
  gc()
  gc()
  
  req(bayes_si_data())
  df <- bayes_si_data()
  
  validate(
    need(nrow(df) >= 5, "insufficient data: you need at least 5 linked pairs.")
  )
  
  stan_data <- list(
    N = nrow(df),
    tStartExposure = df$tStartExposure,
    tEndExposure = df$tEndExposure,
    tSymptomOnset = df$tSymptomOnset
  )
  
  withProgress(message = 'running bayesian model...', detail = 'initializing...', value = 0, {
    
    mod <- get_compiled_stan_model(input$bayes_model_select_si)
    incProgress(0.1, detail = "sampling chains...")
    
    # run the sampling
    fit <- rstan::sampling(
      object = mod, 
      data = stan_data,
      chains = 1, 
      cores = 1, 
      iter = input$iter_count_si, 
      warmup = input$iter_count_si / 2,
      refresh = 0,
      # optimization: only save specific parameters to save ram
      pars = c("alphaInc", "sigmaInc"),
      control = list(adapt_delta = 0.95)
    )
    
    incProgress(0.8, detail = "extracting posteriors...")
    params <- rstan::extract(fit)
    
    # final cleanup after heavy processing
    gc()
    
    list(
      fit_obj = fit,
      iterations = input$iter_count_si,
      model_name = input$bayes_model_select_si,
      alpha = params$alphaInc,
      sigma = params$sigmaInc,
      alpha_summary = quantile(params$alphaInc, probs = c(0.025, 0.5, 0.975)),
      sigma_summary = quantile(params$sigmaInc, probs = c(0.025, 0.5, 0.975))
    )
  })
})

```

## Row {data-height="180"}

### Sample Size

```{r si-bayes-sample-size}

renderUI({
  req(bayes_si_data())
  
  ss <- nrow(bayes_si_data())
  
  htmltools::div(
    style = "display: flex; flex-direction: column; justify-content: center; align-items: center; height: 100%; width: 100%; padding: 10px;",
    htmltools::div(
      style = "font-size: 40px; font-weight: bold; margin-bottom: 5px; text-align: center; font-family: Tw Cen MT; color: #000000;",
      ss
    ),
    htmltools::div(
      style = "font-size: 14px; text-align: center; color: #666;",
      "Linked Pairs"
    )
  )
})

```

### Model Statistics (Median & 95% CrI)

```{r si-bayes-model-stats}

renderDT({
  req(bayes_si_fit())
  res <- bayes_si_fit()
  
  # dynamic parameter labels
  if (res$model_name == "Lognormal") {
    p_names <- c("MeanLog (Alpha)", "SDLog (Sigma)")
    post_means <- exp(res$alpha + (res$sigma^2)/2)
  } else if (res$model_name == "Gamma") {
    p_names <- c("Shape (Alpha)", "Rate (Sigma)")
    post_means <- res$alpha / res$sigma
  } else if (res$model_name == "Normal") {
    p_names <- c("Mean (Alpha)", "SD (Sigma)")
    post_means <- res$alpha # mean is just alpha for normal
  } else {
    p_names <- c("Shape (Alpha)", "Scale (Sigma)")
    post_means <- res$sigma * gamma(1 + 1/res$alpha)
  }
  
  df_stats <- data.frame(
    Parameter = p_names,
    `Median` = round(c(res$alpha_summary[2], res$sigma_summary[2]), 3),
    `Lower_95_CrI` = round(c(res$alpha_summary[1], res$sigma_summary[1]), 3),
    `Upper_95_CrI` = round(c(res$alpha_summary[3], res$sigma_summary[3]), 3)
  )
  
  mean_summ <- quantile(post_means, probs = c(0.025, 0.5, 0.975))
  
  df_mean <- data.frame(
    Parameter = "Mean Serial Interval",
    `Median` = round(mean_summ[2], 3),
    `Lower_95_CrI` = round(mean_summ[1], 3),
    `Upper_95_CrI` = round(mean_summ[3], 3)
  )
  
  final_df <- rbind(df_stats, df_mean)
  
  fname <- paste0("Serial_Interval_Bayesian_Stats_", Sys.Date())
  
  datatable(
    final_df, 
    rownames = FALSE, 
    extensions = 'Buttons',
    options = list(
      dom = 'Brt',
      buttons = list(
        list(extend = "csv", filename = fname),
        list(extend = "excel", filename = fname),
        list(extend = "pdf", filename = fname)
      ),
      columnDefs = list(list(className = 'dt-center', targets = "_all"))
    )
  ) %>%
    formatStyle(columns = "Parameter", fontWeight = "bold")
})

```

Column {.tabset .tabset-fade }
-------------------------------

### Dataset Preview

```{r si-bayes-dataset-preview}

renderDataTable({
  bayes_si_data()
}, options = list(
  pageLength = 10,
  paging = TRUE,
  striped = TRUE,
  bordered = TRUE,
  hover = TRUE
), rownames = FALSE)

```

### Posterior Distribution (Median Fit)

```{r si-bayes-posterior-dist}

renderHighchart({
  req(bayes_si_fit())
  res <- bayes_si_fit()
  
  alpha_est <- median(res$alpha)
  sigma_est <- median(res$sigma)
  
  # dynamic x-axis range logic
  if (res$model_name == "Normal") {
    # calculate lower bound to capture negative tail (99.9% coverage)
    lower_bound <- qnorm(0.001, mean = alpha_est, sd = sigma_est)
    # cap upper bound at 30 for comparability
    upper_bound <- 30
    x_vals <- seq(lower_bound, upper_bound, length.out = 300)
    x_min_limit <- NULL # let chart auto-scale to show negatives
  } else {
    # for strictly positive models, fix range 0 to 30
    x_vals <- seq(0, 30, length.out = 300)
    x_min_limit <- 0
  }
  
  # dynamic density calculation
  if (res$model_name == "Gamma") {
    y_vals <- dgamma(x_vals, shape = alpha_est, rate = sigma_est)
  } else if (res$model_name == "Lognormal") {
    y_vals <- dlnorm(x_vals, meanlog = alpha_est, sdlog = sigma_est)
  } else if (res$model_name == "Normal") {
    y_vals <- dnorm(x_vals, mean = alpha_est, sd = sigma_est)
  } else {
    y_vals <- dweibull(x_vals, shape = alpha_est, scale = sigma_est)
  }
  
  model_label <- tools::toTitleCase(res$model_name)
  menu_items <- list("viewFullscreen", "separator", "downloadPNG", "downloadPDF", "downloadSVG")
  axis_style <- list(color = "#000000", fontSize = "14px", fontWeight = "bold", fontFamily = "Tw Cen MT")
  
  highchart() %>%
    hc_chart(
      style = list(fontFamily = "Tw Cen MT"), 
      backgroundColor = "#FFFFFF"
    ) %>%
    
    hc_title(
      text = paste0("Posterior Predictive Distribution (", model_label, ")"),
      style = list(fontSize = "18px", fontWeight = "bold", fontFamily = "Tw Cen MT", color = "#000000")
    ) %>%
    
    hc_subtitle(
      text = paste0("Based on ", res$iterations, " MCMC iterations"),
      style = list(fontSize = "14px", fontFamily = "Tw Cen MT", color = "#666666")
    ) %>%
    
    hc_exporting(
      enabled = TRUE, 
      filename = paste0("Serial_Interval_Posterior_Distribution_", Sys.Date()),
      buttons = list(contextButton = list(menuItems = menu_items))
    ) %>%
    
    hc_add_series(
      data = data.frame(x = x_vals, y = y_vals), 
      type = "area", 
      name = "Posterior Density", 
      color = "rgba(255, 140, 0, 0.5)",
      lineWidth = 2
    ) %>%
    
    hc_xAxis(
      title = list(text = "Serial Interval (Days)", style = axis_style), 
      labels = list(style = axis_style),
      min = x_min_limit,
      minPadding = 0,
      startOnTick = TRUE,
      lineColor = "#000000", 
      lineWidth = 1, 
      tickColor = "#000000"
    ) %>%
    
    hc_yAxis(
      title = list(text = "Density", style = axis_style), 
      labels = list(style = axis_style),
      min = 0, 
      minPadding = 0, 
      startOnTick = TRUE,
      gridLineWidth = 0, 
      lineColor = "#000000", 
      lineWidth = 1, 
      tickColor = "#000000"
    ) %>%
    
    hc_tooltip(valueDecimals = 4, style = list(fontFamily = "Tw Cen MT"))
})

```

### Trace Plot (Convergence Check)

```{r si-bayes-trace-plot}

renderPlotly({
  req(bayes_si_fit())
  
  p <- rstan::traceplot(bayes_si_fit()$fit_obj, 
                        pars = c("alphaInc", "sigmaInc"),
                        linewidth = 0.25) +
    theme_bw() +
    ggtitle("Trace Plot: Alpha & Sigma") + 
    theme(
      text = element_text(family = "Tw Cen MT"),
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      axis.title = element_text(size = 12, face = "bold"),
      axis.text = element_text(size = 10, color = "black"),
      axis.line = element_line(color = "black"),
      strip.background = element_rect(fill = "#f0f0f0"),
      strip.text = element_text(face = "bold")
    )
  
  ggplotly(p) %>%
    config(
      toImageButtonOptions = list(
        format = "png",
        filename = paste0("Serial_Interval_Trace_Plot_", Sys.Date()),
        height = 500,
        width = 700,
        scale = 2
      )
    ) %>%
    layout(font = list(family = "Tw Cen MT"))
})

```


$k$ and $R_0$ {data-orientation=rows}
=======================================

Sidebar {.sidebar}
-------------------

```{r}

hr()
fileInput("file1_kr", "Choose CSV File",
          multiple = TRUE,
          accept = c("text/csv", "text/comma-separated-values,text/plain", ".csv"))
hr()
# cluster selection for transmission network visualization - initialize with placeholder
selectInput(inputId = "cluster_kr", label = "Select cluster:",
            choices = c("All"), selected = "All")
hr()
valueBox(
  value = tags$h6("Note: Superspreading event is an event in which an infectious disease is spread much more than usual. Superspreader: An unusually contagious individual infected with a disease. The superspreading potential is estimated through the R₀ (Basic Reproduction number) and Kappa (variation in individual infectiousness).", 
                  style = "text-align: justify; font-family: Gill Sans MT; font-size: 16px; color: white;"),
  color = "#FFFFFF"
)
hr()
valueBox(
  value = downloadButton("downloadMockKappa", "Download Mock Data", icon = icon("download")),
  color = "#FFFFFF"
)

# logo image
renderImage({
  list(
    src = "one_logo.png",
    contentType = 'image/png',
    width = 210,
    height = 60,
    style = "position: fixed; bottom: 10px; left: 10px; cursor: pointer;",
    onclick = "window.location.href = 'http://onehealth.or.kr/';"
  )
}, deleteFile = FALSE)

```

```{r k-and-r0-data-processing}

# reactive data for k and r0 - filtered by selected cluster
dataset_kr <- eventReactive(input$file1_kr, {
  req(input$file1_kr)
  dataset <- read.csv(input$file1_kr$datapath)
  dataset <- data.frame(lapply(dataset, trimws))
  dataset <- janitor::clean_names(dataset)
  
  # validation start
  validate(
    need(all(c("cluster_id", "transmission_relationship") %in% names(dataset)),
         "incorrect dataset: please upload k and r0 data containing 'cluster id' and 'transmission relationship'.")
  )
  # validation end
  
  dataset
})

# reactive dataset filtered by selected cluster
filtered_dataset_kr <- reactive({
  req(dataset_kr(), input$cluster_kr)
  dataset <- dataset_kr()
  
  # filter data based on selected cluster
  if (input$cluster_kr != "All") {
    dataset <- dataset %>% filter(cluster_id == input$cluster_kr)
  }
  
  dataset
})

# update cluster choices when data loads
observeEvent(dataset_kr(), {
  req(dataset_kr())
  clusters <- c("All", unique(dataset_kr()$cluster_id) %>% sort())
  updateSelectInput(session, "cluster_kr", choices = clusters, selected = "All")
})

# helper function for offspring distribution - uses filtered data
get_offspring_data <- function() {
  req(filtered_dataset_kr())
  dt <- filtered_dataset_kr()
  if (nrow(dt) == 0) stop("no data available for selected cluster")
  freqt <- table(dt$source_of_infection_id) %>% data.frame() %>% filter(Var1 != 'unknown')
  zeros <- nrow(dt) - nrow(freqt)
  offt <- table(freqt$Freq) %>% data.frame()
  offtk <- c(rep(0, zeros), rep(as.numeric(levels(offt$Var1)), offt$Freq))
  list(offtk = offtk, fit = fitdistr(offtk, "negative binomial"))
}

```

## Row {data-height="180"}

### Sample Size

```{r}

renderUI({
  # ensure filtered dataset is available
  req(filtered_dataset_kr())
  
  # data Preparation
  ss <- nrow(filtered_dataset_kr())
  
  # render the sample size text
  htmltools::div(
    style = "display: flex; flex-direction: column; justify-content: center; align-items: center; height: 100%; width: 100%; padding: 10px;",
    htmltools::div(
      style = "font-size: 30px; font-weight: bold; margin-bottom: 10px; text-align: center;font-family: Tw Cen MT;",
      ss
    ),
    htmltools::div(
      style = "font-size: 14px; text-align: center; color: #666;",
      paste("Cluster:", input$cluster_kr)
    )
  )
})

```

### Model Parameters

```{r}

renderDT({
  req(input$file1_kr)
  
  res <- tryCatch({
    get_offspring_data()
  }, error = function(e) {
    NULL
  })
  
  if (is.null(res) || is.null(res$fit)) {
    df <- data.frame(
      parameter = c("R₀", "Kappa"),
      value = rep("No data available", 2),
      check.names = FALSE  
    )
  } else {
    fit_df <- as.data.frame(unlist(res$fit))
    fit_df$parameter <- row.names(fit_df)
    rownames(fit_df) <- NULL
    colnames(fit_df) <- c("value", "parameter")
    
    fit_df <- dplyr::select(fit_df, parameter, value) %>%
      mutate(
        parameter = gsub("size", "Kappa", parameter),
        parameter = gsub("mu", "R₀", parameter),
        value = round(value, 3)
      ) %>%
      head(4)
    
    df <- fit_df
  }
  
  datatable(
    df,
    rownames = FALSE,
    options = list(
      dom = "Brt",
      buttons = list(
        list(extend = "csv", filename = paste0("k_ro_model_params_", input$cluster_kr, "_", Sys.Date())),
        list(extend = "excel", filename = paste0("k_ro_model_params_", input$cluster_kr, "_", Sys.Date())),
        list(extend = "pdf", filename = paste0("k_ro_model_params_", input$cluster_kr, "_", Sys.Date()))
      ),
      pageLength = nrow(df)
    ),
    extensions = "Buttons",
    class = "compact cell-border stripe"
  ) %>%
    formatStyle(
      columns = names(df),
      `font-size` = "14px"
    )
})

```

### Confidence Interval

```{r}

renderDT({
  req(input$file1_kr)
  res <- tryCatch({
    get_offspring_data()
  }, error = function(e) {
    NULL
  })
  
  if (is.null(res) || is.null(res$fit)) {
    df <- data.frame(
      parameter = c("R₀", "Kappa"),
      `2.5 %` = rep("No data available", 2),
      `97.5 %` = rep("No data available", 2),
      check.names = FALSE  
    )
  } else {
    conf_interval <- confint(res$fit, level = 0.95)
    conf_df <- as.data.frame(conf_interval)
    conf_df$parameter <- rownames(conf_df)
    
    conf_df <- dplyr::select(conf_df, parameter, `2.5 %`, `97.5 %`) %>%
      mutate(
        parameter = gsub("size", "Kappa", parameter),
        parameter = gsub("mu", "R₀", parameter),
        `2.5 %` = round(`2.5 %`, 3),
        `97.5 %` = round(`97.5 %`, 3)
      )
    
    df <- conf_df
  }
  
  datatable(
    df,
    rownames = FALSE,
    options = list(
      dom = "Brt",
      buttons = list(
        list(extend = "csv", filename = paste0("k_ro_conf_int_", input$cluster_kr, "_", Sys.Date())),
        list(extend = "excel", filename = paste0("k_ro_conf_int_", input$cluster_kr, "_", Sys.Date())),
        list(extend = "pdf", filename = paste0("k_ro_conf_int_", input$cluster_kr, "_", Sys.Date()))
      ),
      pageLength = nrow(df)
    ),
    extensions = "Buttons",
    class = "compact cell-border stripe"
  ) %>%
    formatStyle(
      columns = names(df),
      `font-size` = "14px"
    )
})

```


Column {.tabset .tabset-fade }
-------------------------------

### Histogram

```{r}

renderHighchart({
  
  req(input$file1_kr)
  
  res <- tryCatch({
    get_offspring_data()
  }, error = function(e) {
    NULL
  })
  
  if (is.null(res)) {
    # create empty chart with message
    highchart() %>%
      hc_title(
        text = paste("Only one record available or No data available for cluster:", input$cluster_kr), 
        style = list(
          fontSize = "20px", 
          fontWeight = "bold", 
          color = "#000000", 
          fontFamily = "Tw Cen MT"
        )
      ) %>%
      hc_chart(style = list(fontFamily = "Tw Cen MT"), backgroundColor = "#FFFFFF")
  } else {
    # calculate negative binomial distribution
    xvalue <- 0:max(res$offtk)
    yvalue <- dnbinom(xvalue, size = res$fit$estimate["size"], mu = res$fit$estimate["mu"])
    
    # create histogram data
    hist_data <- hist(res$offtk, plot = FALSE, breaks = seq(-0.5, max(res$offtk) + 0.5, by = 1))
    
    # create data for columns
    hist_df <- data.frame(
      from = hist_data$breaks[-length(hist_data$breaks)],
      to = hist_data$breaks[-1],
      y = hist_data$density,
      x = hist_data$mids  # For positioning reference
    )
    
    # style definitions
    axis_style <- list(color = "#000000", fontSize = "14px", fontWeight = "bold")
    label_style <- list(color = "#000000", fontSize = "16px", fontWeight = "Normal")
    
    # create negative binomial curve data (nudged by -0.5)
    fit_data <- data.frame(x = xvalue - 0.5, y = yvalue)
    
    highchart() %>%
      # add histogram bars
      hc_add_series(
        data = list_parse(hist_df),
        type = "column",
        name = "Histogram",
        color = "#F2F2F2",
        backgroundColor = "#FFFFFF",
        borderColor = "#000000",
        borderWidth = 1,
        groupPadding = -0.05,
        pointPadding = 0.05,
        pointRange = 1,
        pointPlacement = 0
      ) %>%
      # add negative binomial curve
      hc_add_series(
        data = fit_data,
        type = "line",
        name = "Negative Binomial Fit",
        color = "#000000",
        lineWidth = 1,
        marker = list(enabled = FALSE)
      ) %>%
      hc_xAxis(
        title = list(text = "Offspring", style = axis_style),
        allowDecimals = FALSE,
        gridLineWidth = 0, 
        lineColor = "#000000", lineWidth = 1, tickWidth = 1, tickColor = "#000000",
        labels = list(style = label_style),
        min = -0.5,
        max = max(res$offtk) + 0.5,
        tickInterval = 1
      ) %>%
      hc_yAxis(
        title = list(text = "Density", style = axis_style),
        gridLineWidth = 0, 
        lineColor = "#000000", lineWidth = 1, tickWidth = 1, tickColor = "#000000",
        labels = list(style = label_style),
        allowDecimals = TRUE,
        min = 0
      ) %>%
      hc_title(
        text = paste("Cluster:", input$cluster_kr),
        style = list(fontSize = "16px", fontFamily = "Tw Cen MT", fontWeight = "bold")
      ) %>%
      hc_tooltip(
        shared = TRUE,
        crosshairs = TRUE,
        valueDecimals = 2,
        headerFormat = "<b>Offspring: {point.x}</b><br/>",
        pointFormat = "<span style='color:#000000'>Density</span>: <b>{point.y:.2f}</b><br/>"
      ) %>%
      hc_exporting(
        enabled = TRUE,
        filename = paste0("k_ro_", input$cluster_kr),
        buttons = list(
          contextButton = list(
            menuItems = list('downloadPNG', 'downloadPDF', 'downloadSVG')
          )
        )
      ) %>%
      hc_plotOptions(
        column = list(
          grouping = FALSE,
          pointRange = 1,
          pointPlacement = 0
        )
      ) %>%
      hc_legend(enabled = FALSE) %>%
      hc_chart(style = list(fontFamily = "Tw Cen MT"))
  }
})

```

### Transmission relationship

```{r transmission-relationship-output}
visNetworkOutput("network_plot")

```

```{r parse-network-data}

output$download_high_quality <- downloadHandler(
  filename = function() { paste("network_cluster_", input$cluster_kr, ".png", sep = "") },
  content = function(file) {
    temp_html <- tempfile(fileext = ".html")
    visSave(network_plot(), temp_html)
    webshot2::webshot(temp_html, file, vwidth = 3600, vheight = 2800, zoom = 6, delay = 3)
  }
)

# add a new download handler for PDF
output$download_high_quality_pdf <- downloadHandler(
  filename = function() { paste("network_cluster_", input$cluster_kr, ".pdf", sep = "") },
  content = function(file) {
    temp_html <- tempfile(fileext = ".html")
    visSave(network_plot(), temp_html)
    webshot2::webshot(temp_html, file, vwidth = 3600, vheight = 2800, delay = 3)
  }
)

filtered_data <- reactive({
  req(dataset_kr(), input$cluster_kr)
  df <- dataset_kr()
  
  # For transmission relationship, don't filter when "All" is selected
  # as we want to show specific cluster networks
  if (input$cluster_kr != "All") {
    df_filtered <- df %>% filter(cluster_id == input$cluster_kr)
  } else {
    # return empty data frame when "All" is selected
    df_filtered <- df[0, ]  # empty data frame with same structure
  }
  df_filtered
})

network_data <- reactive({
  req(filtered_data())
  df_filtered <- filtered_data()
  
  # Check if data is empty (when "All" is selected)
  if (nrow(df_filtered) == 0) {
    return(list(nodes = data.frame(), edges = data.frame()))
  }
  
  edges_level1_2 <- df_filtered %>%
    dplyr::select(cluster_id, infecteegen1) %>%
    unique() %>%
    rename(from = cluster_id, to = infecteegen1)
  
  edges_level2_3 <- df_filtered %>%
    dplyr::select(infecteegen1, infecteegen2) %>%
    unique() %>%
    rename(from = infecteegen1, to = infecteegen2)
  
  edges_level3_4 <- df_filtered %>%
    dplyr::select(infecteegen2, infecteegen3) %>%
    unique() %>%
    rename(from = infecteegen2, to = infecteegen3)
  
  edges_level4_5 <- df_filtered %>%
    dplyr::select(infecteegen3, infecteegen4) %>%
    unique() %>%
    rename(from = infecteegen3, to = infecteegen4)
  
  edge_list <- rbind(edges_level1_2, edges_level2_3, edges_level3_4, edges_level4_5) %>%
    filter(!from == "" & !to == "" & !is.na(from) & !is.na(to))
  
  all_nodes <- unique(c(edge_list$from, edge_list$to))
  
  node_clusters <- df_filtered %>%
    dplyr::select(id = cluster_id, cluster_id) %>%
    bind_rows(df_filtered %>% dplyr::select(id = infecteegen1, cluster_id)) %>%
    bind_rows(df_filtered %>% dplyr::select(id = infecteegen2, cluster_id)) %>%
    bind_rows(df_filtered %>% dplyr::select(id = infecteegen3, cluster_id)) %>%
    bind_rows(df_filtered %>% dplyr::select(id = infecteegen4, cluster_id)) %>%
    filter(!is.na(id), id != "") %>%
    distinct(id, .keep_all = TRUE)
  
  nodes <- data.frame(
    id = all_nodes,
    label = all_nodes,
    title = paste("ID:", all_nodes, "<br>Cluster:", input$cluster_kr),
    color = "#4169E1",
    font = list(color = "white", face = "bold", size = 40),
    shadow = TRUE,
    size = 25
  ) %>%
    left_join(node_clusters, by = "id") %>%
    mutate(group = ifelse(is.na(cluster_id), "Unknown", as.character(cluster_id)))
  
  list(nodes = nodes, edges = edge_list)
})

output$network_plot <- renderVisNetwork({
  req(network_data())
  data <- network_data()
  nodes <- data$nodes
  edge_list <- data$edges
  
  # Check if "All" is selected or no data available
  if (input$cluster_kr == "All" || nrow(edge_list) == 0) {
    nodes <- data.frame(
      id = 1,
      label = "Please select an individual cluster.",
      title = "Transmission relationships are shown for individual clusters only.<br>Select a specific cluster from the dropdown or check the 'Transmission Networks' tab to see the full network.",
      color = "#f0f0f0",
      font = list(color = "black", size = 20, face = "normal")
    )
    return(
      visNetwork(nodes, data.frame(from = numeric(0), to = numeric(0))) %>%
        visOptions(manipulation = FALSE) %>%
        visNodes(shape = "box", size = 80) %>%
        visLayout(randomSeed = 123) %>%
        visPhysics(stabilization = FALSE) %>%
        visInteraction(dragNodes = FALSE, dragView = FALSE, zoomView = FALSE)
    )
  }
  
  visNetwork(nodes, edge_list) %>%
    visEdges(
      arrows = list(to = list(enabled = TRUE, scaleFactor = 1.2)),
      color = list(color = "#6E6E6E", highlight = "#FF0000"),
      smooth = list(enabled = TRUE, type = "dynamic")
    ) %>%
    visNodes(
      shape = "circle",
      size = 90,
      borderWidth = 1,
      borderWidthSelected = 4,
      font = list(size = 24, color = "white", face = "bold")
    ) %>%
    visOptions(
      highlightNearest = list(enabled = TRUE, degree = list(from = 5, to = 5), hover = TRUE, algorithm = "hierarchical"),
      nodesIdSelection = list(enabled = TRUE, values = nodes$id, selected = input$cluster_kr, style = 'width: 200px; height: 26px;')
    ) %>%
    visExport(
      type = "png",
      name = paste("transmission_network_cluster", input$cluster_kr),
      label = "Download as PNG",
      background = "#fff",
      style = "position: absolute; top: -400px; right: 10px; z-index: 10;",
      loadDependencies = TRUE
    ) %>%
    visInteraction(
      navigationButtons = TRUE,
      keyboard = TRUE,
      tooltipDelay = 0,
      dragView = TRUE,
      zoomView = TRUE
    ) %>%
    visLayout(
      randomSeed = 123,
      improvedLayout = TRUE
    ) %>%
    visPhysics(
      stabilization = list(iterations = 100),
      solver = "forceAtlas2Based",
      forceAtlas2Based = list(
        gravitationalConstant = -50,
        centralGravity = 0.01,
        springLength = 100,
        springConstant = 0.08
      )
    ) %>%
    visLegend(enabled = FALSE) %>%
    visEvents(
      selectNode = "function(nodes) { Shiny.setInputValue('selected_node', nodes); }"
    )
})

observeEvent(input$selected_node, {
  req(input$selected_node, dataset_kr())
  selected_id <- input$selected_node$nodes[[1]]
  df <- dataset_kr()
  selected_cluster <- df %>%
    filter(cluster_id == selected_id |
             infecteegen1 == selected_id |
             infecteegen2 == selected_id |
             infecteegen3 == selected_id |
             infecteegen4 == selected_id) %>%
    pull(cluster_id) %>%
    first()
  if(!is.na(selected_cluster) && selected_cluster != input$cluster_kr) {
    updateSelectInput(session, "cluster_kr", selected = selected_cluster)
  }
})

```

### Transmission Networks

```{r transmission-networks-output}
visNetworkOutput("full_network_plot")
```


```{r parse-transmission-chain}

# function to parse transmission chains
parse_transmission_chain <- function(relationship) {
  # skip if relationship is empty or unknown
  if (is.na(relationship) || relationship == "" || relationship == "unknown") {
    return(data.frame(from = character(), to = character()))
  }
  # split the chain by the arrow symbol (noticed this â†’ in the csv)
  chain <- strsplit(relationship, "→|â†’")[[1]]
  chain <- trimws(chain)  # added this cos some rows have whitespace
  
  # extract immediate relationships (each consecutive pair)
  if (length(chain) >= 2) {
    edges <- data.frame(
      from = chain[1:(length(chain) - 1)],
      to = chain[2:length(chain)]
    )
    return(edges[edges$from != "" & edges$to != "" & edges$from != edges$to, ])
  }
  
  return(data.frame(from = character(), to = character()))
}

output$download_full_network <- downloadHandler(
  filename = function() { paste("full_transmission_network_", Sys.Date(), ".png", sep = "") },
  content = function(file) {
    temp_html <- tempfile(fileext = ".html")
    visSave(full_network_plot(), temp_html)
    webshot2::webshot(temp_html, file, vwidth = 3600, vheight = 2800, zoom = 6, delay = 3)
  }
)

full_network_data <- reactive({
  req(dataset_kr())
  df <- dataset_kr()
  
  edges_list <- lapply(df$transmission_relationship, parse_transmission_chain)
  edges <- do.call(rbind, edges_list)
  
  edges <- edges[complete.cases(edges), ]
  edges <- unique(edges)
  
  # get all unique nodes and sort them numerically
  all_nodes <- unique(c(edges$from, edges$to))
  
  # extract numeric values for sorting
  node_numbers <- as.numeric(gsub("#", "", all_nodes))
  sorted_nodes <- all_nodes[order(node_numbers)]
  
  # create nodes data frame
  nodes <- data.frame(
    id = sorted_nodes,
    label = sorted_nodes,
    title = paste("ID:", sorted_nodes),
    color = "#4169E1",
    font = list(color = "white", face = "bold", size = 50),
    shadow = TRUE,
    size = 30
  )
  
  list(nodes = nodes, edges = edges, sorted_ids = sorted_nodes)
})

output$full_network_plot <- renderVisNetwork({
  req(full_network_data())
  data <- full_network_data()
  nodes <- data$nodes
  edges <- data$edges
  sorted_ids <- data$sorted_ids
  
  if(nrow(edges) == 0) {
    nodes <- data.frame(
      id = 1,
      label = "No transmission data",
      title = "No transmission relationships found in the dataset",
      color = "#f0f0f0",
      font = list(color = "black", size = 24)
    )
    return(
      visNetwork(nodes, data.frame(from = numeric(0), to = numeric(0))) %>%
        visOptions(manipulation = FALSE) %>%
        visNodes(shape = "box", size = 50) %>%
        visLayout(randomSeed = 123)
    )
  }
  
  visNetwork(nodes, edges) %>%
    visEdges(
      arrows = list(to = list(enabled = TRUE, scaleFactor = 1.2)),
      color = list(color = "#6E6E6E", highlight = "#FF0000"),
      smooth = list(enabled = TRUE, type = "dynamic")
    ) %>%
    visNodes(
      shape = "circle",
      size = 30,
      borderWidth = 1,
      borderWidthSelected = 4,
      font = list(size = 20, color = "white", face = "bold")
    ) %>%
    visOptions(
      highlightNearest = list(enabled = TRUE, degree = 1, hover = TRUE),
      nodesIdSelection = list(
        enabled = TRUE, 
        values = sorted_ids,
        style = 'width: 200px; height: 26px;'
      )
    ) %>%
    visExport(
      type = "png",
      name = paste("full_transmission_network_", Sys.Date()),
      label = "Download as PNG",
      background = "#fff",
      style = "position: absolute; top: -400px; right: 10px;",
      loadDependencies = TRUE
    ) %>%
    visInteraction(
      navigationButtons = TRUE,
      keyboard = TRUE,
      tooltipDelay = 0,
      dragView = TRUE,
      zoomView = TRUE
    ) %>%
    visLayout(
      randomSeed = 123,
      improvedLayout = TRUE
    ) %>%
    visPhysics(
      stabilization = list(iterations = 100),
      solver = "forceAtlas2Based",
      forceAtlas2Based = list(
        gravitationalConstant = -50,
        centralGravity = 0.01,
        springLength = 100,
        springConstant = 0.08
      )
    ) %>%
    visLegend(enabled = FALSE)
})

```

### Dataset Preview

```{r kappa-r0-dataset-preview}
renderDataTable({
  filtered_dataset_kr()
}, options = list(pageLength = 10), rownames = FALSE)

```
